---
title: "Binary Analysis"
author: Simmons & Girard
date: April 15, 2026
theme: lumen
format:
  html:
    embed-resources: true
---


# 1. Prep Data

## 1.1. Libraries
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(simhelpers)
library(knitr)
library(kableExtra)
```

## 1.2. Directory & Loading Data

```{r}
DIR <- "~/Documents/Github/MastersThesis/iccs/data"
dat_nest <- readRDS(paste0(DIR,'/binary_results_0312_01.rds'))
dat_aov_res <- readRDS(paste0(DIR,'/binary_aOV_full.rds')) #aov results

# dat_all <-  dat_nest |> left_join(dat_aov, by = c("n_raters", "n_objects","target_icc","p","seed"))

#head(dat_nest) 

#Kable (doesn't render?)
# k1 <- kable(dat_nest, format = "html") %>%
#   kable_styling(bootstrap_options = c("striped","hover")) %>%
#   scroll_box(width = "100%", height = "200%")
# k1 

```


## 1.3. Data Wrangling 
```{r}
## unnest the results' list-column  (by hand; example with "icc" column which is numeric) 

# dat <- dat_nest |> 
#   mutate(result = map(result, as_tibble)) |>
#   unnest(result) |>
#   mutate(
#     icc = map_dbl(icc, \(x) pluck(x, .default= NA_real_))) #example

##  unnest efficiently, let R decide the best atomic data type
dat_all <- dat_nest |> 
  mutate(result = map(result, as_tibble)) |>
  unnest(result) |>
  #remove columns for now (unnecessary)
  select(!c(approach, observed, expected, adjusted, boot_results, details, call)) |>
  mutate(across(where(is.list), \(col) {
    map_vec(col, \(x) if(is.null(x)) NA else x)
  }))

dat_aov <- dat_aov_res |>
  mutate(result = map(result, as_tibble)) |>
  unnest(result) |>
  #remove columns for now (unnecessary)
 # select(!c(approach, observed, expected, adjusted, boot_results, details, call)) |>
  mutate(across(where(is.list), \(col) {
    map_vec(col, \(x) if(is.null(x)) NA else x)
  }))

# dat <- dat_all |> 
#   left_join(dat_aov, 
#   by = c("n_raters", "n_objects","target_icc","p","seed","filename"))

#set conditions as factors
conds <- c("n_raters", "n_objects", "p")
dat_all <- dat_all |> 
  mutate(across(.cols = all_of(conds), .fns = as.factor))

dat_aov <- dat_aov |> 
  mutate(across(.cols = all_of(conds), .fns = as.factor))



# Delete nested object
rm(dat_nest)
rm(dat_aov_res)

# Print
head(dat_all)
head(dat_aov)

```

<b> In the binary case, the parameter grid of (54) conditions is as follows: </b>

```{r}
design_factors <- list(
n_raters = c(3,12,24),
n_objects = c(10,50,100),
target_icc = c(0.40,0.60,0.80),
p = c(0.5, 0.8)# for the binary case
)  
#expand grid
params <- expand_grid( !!!design_factors) 
design_factors
head(params)
```

# 2. Performance Measures for Point Estimators
Here, we consider measures for point estimators for each of the 54 binary conditions

  - Bias and Standard Errors in Estimators

For all measures 
```{r}
bias_estimators_all <- dat_all |>
  #group by each condition (For ordinal, 162 total conditions)
  select(!c(seed, filename)) |>
  group_by(n_raters, n_objects, target_icc, p) |> 
  summarize(
    mean_icc_hat = mean(icc, na.rm = TRUE), #across 1000 iterations of simulation
    se_icc_hat = sd(icc, na.rm = TRUE), # SD of sampling dist. is SE
    dropped_icc = sum(is.na(icc))/sum(!is.na(icc)) *100,
    mean_gICC_hat = mean(g_icc, na.rm = TRUE),#across 1000 iterations of simulation
    se_gICC_hat = sd(g_icc, na.rm = TRUE), #SD of sampling dist. is SE
    dropped_gICC = sum(is.na(g_icc))/sum(!is.na(g_icc)) * 100) |>
  # let's number each condition/case for simplicity
  ungroup()|>
  mutate(CaseNumber = row_number()) |>
  mutate(
    target_icc = as.numeric(target_icc),
    bias_icc = mean_icc_hat - target_icc,
    #compare gICC to benchmark of generalized ICC of same condition
    #this assumes the gICC is more accurate than the ICC 
    per_SE_icc_hat = se_icc_hat / se_gICC_hat, 
    bias_gICC = mean_gICC_hat - target_icc,
    per_SE_gICC_hat = se_gICC_hat/se_gICC_hat)
    
head(bias_estimators_all)
```

For AOV measures 

```{r}
bias_estimators_aov <- dat_aov |>
  #group by each condition (For ordinal, 162 total conditions)
  select(!c(seed, filename)) |>
  group_by(n_raters, n_objects, target_icc, p) |> 
  summarize(
    mean_icc_aov = mean(aov_icc, na.rm = TRUE), #across 1000 iterations of simulation
    se_icc_aov = sd(aov_icc, na.rm = TRUE), # SD of sampling dist. is SE
    dropped_icc = sum(is.na(aov_icc))/sum(!is.na(aov_icc)) *100) |>
  # let's number each condition/case for simplicity
  ungroup()|>
  mutate(CaseNumber = row_number()) |>
  mutate(
    target_icc = as.numeric(target_icc),
    bias_aov = mean_icc_aov - target_icc)
    
head(bias_estimators_aov)
```

Join together, sharing CaseNumber
```{r}

bias_all <-  bias_estimators_all |> left_join(bias_estimators_aov, by = c("n_raters", "n_objects","target_icc","p","CaseNumber"))
```
## 2.1 Discussion: Point Estimates Bias
It appears that the traditional ICC (using \{lme4\}) is systematically negatively biased in all conditions from the true ICC value from -23% (Case #5) to -13% (Case #7) in ICC units, all when n_raters = 3! 
```{r}
# range(bias_estimators$bias_icc)
# rbind(bias_estimators[5,], #-0.23
#       bias_estimators[7,]) #-0.13 
```

It appears that the generalized ICC (using \{glmmTMB\}) is biased from -7.87% to -0.004% (in ICC units), but note that these are in cases when there is only 3 raters! 

```{r}
# range(bias_estimators$bias_gICC)
# rbind(bias_estimators[5,], #-0.078
#       bias_estimators[50,]) #-0.0035
```

# Plotting Bias 

```{r}
bias_long_p0.5 <- bias_all |>
  pivot_longer(
    cols = starts_with("bias"),
    names_prefix = "bias_",
    names_to = "method",
    values_to = "bias" 
  ) |> filter(p == 0.5)

b1 <- ggplot(bias_long_p0.5, aes(n_raters, bias, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  geom_point() + geom_line() + 
  geom_hline( yintercept = 0 ) +
  theme_minimal() + ggtitle("Binary, p =0.5")

bias_long_p0.8 <- bias_all |>
  pivot_longer(
    cols = starts_with("bias"),
    names_prefix = "bias_",
    names_to = "method",
    values_to = "bias" 
  ) |> filter(p == 0.8)

b2 <- ggplot(bias_long_p0.8, aes(n_raters, bias, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  geom_point() + geom_line() + 
  geom_hline( yintercept = 0 ) +
  theme_minimal() + ggtitle("Binary, p =0.8")
```


