---
title: "Ordinal Analysis (Universe #2)"
author: Simmons & Girard
date: March 18, 2026
theme: lumen
format:
  html:
    embed-resources: true
---


# 1. Prep Data

## 1.1. Libraries
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(simhelpers)
library(knitr)
library(kableExtra)
```
## 1.2. Directory & Loading Data

```{r}
DIR <-  "~/Documents/Github/MastersThesis/iccs/data/universe_2/"

nest_rds_files <- function(data_df, filename_col, directory_path) {
  # 1. Ensure the directory path ends with a slash for file path construction
  dir <- ifelse(grepl("/$", directory_path), directory_path, paste0(directory_path, "/"))
  
  # 2. Create the new nested column
  # We use sym() to handle the column name if passed as a string
  result_df <- data_df %>%
    mutate(result = map(!!sym(filename_col), function(fname) {
      file_path <- paste0(dir, fname)
      
      # Check if file exists to avoid crashing the whole loop
      if (file.exists(file_path)) {
        return(readRDS(file_path))
      } else {
        warning(paste("File not found:", file_path))
        return(NULL)
      }
    }))
  
  return(result_df)
}
```

## 1.3. Data Wrangling 
<b> In the ordinal case, the parameter grid of conditions is as follows and add each file to its condition </b>

```{r}
# Join each data file to its condition 
design_factors <- list(
n_raters = c(3,12,24),
n_objects = c(10,50,100),
target_icc = c(0.40,0.60,0.80),
k_category = c(3,5,7),
e_category = c(TRUE,FALSE) # 1 = equal category prevalence, #0 = unequal linear decay prevalence 
)
#expand grid
params <- expand_grid( !!!design_factors) |>
    mutate(
    seed = 03122026 + 17 * 1:n(), #set seed for each row,
    condition = 1:n() * 1,
    filename = paste0("ordinal_",condition,"_",seed,".rds")
  )
head(params)

dat_nest <- nest_rds_files(params, "filename", DIR)

# load the separate AOV file 
dat_aov <- readRDS("~//Documents/Github/MastersThesis/iccs/data/ordinal_aov.rds")

```

```{r}


## unnest the results' list-column  (by hand; example with "icc" column which is numeric) 

dat_all <- dat_nest |> unnest(result)
dat_aov <- dat_aov |> unnest(result)

# create full dataset 
dat <- dplyr::bind_rows(dat_all,dat_aov)

# Delete nested object
rm(dat_nest)
rm(dat_aov)
rm(dat_all)

# Print
head(dat)

#set conditions as factors 
conds <- c("n_raters", "n_objects", "k_category", "icc", "e_category", "condition", "method")
dat <- dat |> 
  mutate(across(.cols = all_of(conds), .fns = as.factor))

```


# 2. Performance Measures for Point Estimators
Here, we consider performance measures for point estimators for each of the 162 ordinal conditions. 

We obtain, for each condition and measure (ICC, Generalized ICC, and AOV_ICC):
- Bias (defined as "Mean absolute error" by Oest & Girard, Miratrix & Pustejovsky)
- Bias (defined by Muthen & Muthen)
- Standard Errors (i.e., standard deviation of replication distribution) 
- Root Mean Squared Error (RMSE) 

```{r}
bias_full <- dat |>
  #group by each condition (For ordinal, 162 total conditions)
  select(!c(seed, filename)) |>
  drop_na() |> 
  mutate(estimate = na_if(estimate,NaN)) |> 
 # filter(icc== "ICC(A,1)") |> 
 # group_by(n_raters, n_objects, target_icc, k_category, e_category) |> 
  summarize(
    simhelpers::calc_absolute( 
      estimates = estimate, true_param = target_icc, 
      criteria = c("bias","stddev","rmse")
    ), 
    .by = c(n_raters, n_objects, target_icc, icc, k_category, e_category, condition, method)
  )
    
head(bias_full)
```


## 2.1 Discussion: Point Estimates Bias
Note: all graphs have corresponding Monte Carlo Standard Error confidence intervals!

### 2.1.1 Bias (Mean Absolute Error)

vibe coding 

```{r}

my_geom <- list(
  geom_point(position = position_dodge(width = 0.3)),
  geom_line(position = position_dodge(width = 0.3)), 
  geom_errorbar( aes( ymin = bias - 1.96*bias_mcse,
                      ymax = bias + 1.96*bias_mcse ),
                 position = position_dodge( width=0.3),
                 width = 0 ), 
  geom_hline( yintercept = 0 ),
  theme_minimal() 
)
plot_simulation_bias <- function(data, x_var, y_var, col_var, facet_row, facet_col, title = "Simulation Bias") {
  
  ggplot(data, aes(x = {{x_var}}, 
                   y = {{y_var}}, 
                   col = {{col_var}}, 
                   pch = {{col_var}}, 
                   group = {{col_var}})) +
    # facet_grid handles the two factors dynamically
    facet_grid(rows = vars({{facet_row}}), 
               cols = vars({{facet_col}}), 
               labeller = label_both) +
    # your custom geom layer
    my_geom + 
    labs(title = title,
         x = substitute(x_var),
         y = substitute(y_var),
         color = substitute(col_var),
         shape = substitute(col_var)) +
    theme_bw() # Or your preferred base theme
}

plot_list <- bias_full %>%
  group_by(k_category, e_category) %>%
  group_split()

# 2. Map the plotting function over the list
# We use 'map' to generate a list of ggplot objects
all_plots <- bias_full %>%
  group_by(icc, k_category, e_category) %>%
  group_split() %>%
  # 2. Map over the split data
  map(~ {
    # Extract values for a dynamic and descriptive title
    icc_val <- unique(.x$icc)
    k_val   <- unique(.x$k_category)
    e_val   <- unique(.x$e_category)
    
    # Generate the plot using your generalized function
    plot_simulation_bias(
      data = .x,
      x_var = n_raters,
      y_var = bias,
      col_var = method,
      facet_row = target_icc,
      facet_col = n_objects,
      title = paste0("ICC Type: ", icc_val, " | K = ", k_val, ", Equal = ", e_val)
    )
  })
```



### 2.1.2 Bias (Muthen & Muthen)
- Here, if bias is less than 10% then it is good
```{r}
my_geom2 <- list(
  geom_point(position = position_dodge(width = 0.3)),
  geom_line(position = position_dodge(width = 0.3)), 
  geom_errorbar( aes( ymin = bias2 - 1.96*bias_mcse,
                      ymax = bias2 + 1.96*bias_mcse ),
                 position = position_dodge( width=0.3),
                 width = 0 ), 
  geom_hline( yintercept = 0 ),
  theme_minimal(), 
  geom_hline(yintercept = -0.10, linetype='dotted', col = 'red'),
  scale_y_continuous(labels = scales::label_percent())
)

bias_long_k3_equal <- bias_full |>
  mutate(bias2 = bias/target_icc) |>
  filter(k_category == 3 & e_category == "TRUE") 

p7 <- ggplot(bias_long_k3_equal, aes(n_raters, bias2, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K = 3, equal") + 
  my_geom2

bias_long_k3_notequal <- bias_full |>
  mutate(bias2 = bias/target_icc) |>
 filter(k_category == 3 & e_category == "FALSE") 

p8 <- ggplot(bias_long_k3_notequal, aes(n_raters, bias2, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
 ggtitle("K=3, not equal") + 
 my_geom2

bias_long_k5_equal <- bias_full |>
  mutate(bias2 = bias/target_icc) |>
filter(k_category == 5 & e_category == "TRUE") 

p9 <- ggplot(bias_long_k5_equal, aes(n_raters, bias2, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
 ggtitle("K = 5, equal") +
 my_geom2

bias_long_k5_notequal <- bias_full |>
  mutate(bias2 = bias/target_icc) |>
 filter(k_category == 5 & e_category == "FALSE") 


p10 <- ggplot(bias_long_k5_notequal, aes(n_raters, bias2, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K=5, not equal") +
  my_geom2

bias_long_k7_equal <- bias_full |>
  mutate(bias2 = bias/target_icc) |>
 filter(k_category == 7 & e_category == "TRUE") 

p11 <- ggplot(bias_long_k7_equal, aes(n_raters, bias2, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K=7, equal") +
  my_geom2

bias_long_k7_notequal <- bias_full |>
  mutate(bias2 = bias/target_icc) |>
 filter(k_category == 7 & e_category == "FALSE") 

p12 <- ggplot(bias_long_k7_notequal, aes(n_raters, bias2, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K=7, not equal") + 
  my_geom2

p7
p8
p9
p10
p11
p12

#### ICC (A,1) MAE Bias results 
```{r}

plot_list <- bias_ICC_A_1 %>%
  group_by(k_category, e_category) %>%
  group_split()

# 2. Map the plotting function over the list
# We use 'map' to generate a list of ggplot objects
A_1_plots <- bias_ICC_A_1 %>%
  group_by(icc, k_category, e_category) %>%
  group_split() %>%
  # 2. Map over the split data
  map(~ {
    # Extract values for a dynamic and descriptive title
    icc_val <- unique(.x$icc)
    k_val   <- unique(.x$k_category)
    e_val   <- unique(.x$e_category)
    
    # Generate the plot using your generalized function
    plot_simulation_bias(
      data = .x,
      x_var = n_raters,
      y_var = bias,
      col_var = method,
      facet_row = target_icc,
      facet_col = n_objects,
      title = paste0("ICC Type: ", icc_val, " | K = ", k_val, ", Equal = ", e_val)
    )
  })
```

#### ICC (A,k) MAE Bias results 

```{r}

plot_list <- bias_ICC_A_K %>%
  group_by(k_category, e_category) %>%
  group_split()

# 2. Map the plotting function over the list
# We use 'map' to generate a list of ggplot objects
A_k_plots <- bias_ICC_A_K %>%
  group_by(icc, k_category, e_category) %>%
  group_split() %>%
  # 2. Map over the split data
  map(~ {
    # Extract values for a dynamic and descriptive title
    icc_val <- unique(.x$icc)
    k_val   <- unique(.x$k_category)
    e_val   <- unique(.x$e_category)
    
    # Generate the plot using your generalized function
    plot_simulation_bias(
      data = .x,
      x_var = n_raters,
      y_var = bias,
      col_var = method,
      facet_row = target_icc,
      facet_col = n_objects,
      title = paste0("ICC Type: ", icc_val, " | K = ", k_val, ", Equal = ", e_val)
    )
  })
```

#### ICC(C,1) MAE bias results
```{r}

plot_list <- bias_ICC_C_1 %>%
  group_by(k_category, e_category) %>%
  group_split()

# 2. Map the plotting function over the list
# We use 'map' to generate a list of ggplot objects
C_1_plots <- bias_ICC_C_1 %>%
  group_by(icc, k_category, e_category) %>%
  group_split() %>%
  # 2. Map over the split data
  map(~ {
    # Extract values for a dynamic and descriptive title
    icc_val <- unique(.x$icc)
    k_val   <- unique(.x$k_category)
    e_val   <- unique(.x$e_category)
    
    # Generate the plot using your generalized function
    plot_simulation_bias(
      data = .x,
      x_var = n_raters,
      y_var = bias,
      col_var = method,
      facet_row = target_icc,
      facet_col = n_objects,
      title = paste0("ICC Type: ", icc_val, " | K = ", k_val, ", Equal = ", e_val)
    )
  })
```

#### (C,K) bias results 
```{r}

plot_list <- bias_ICC_C_K %>%
  group_by(k_category, e_category) %>%
  group_split()

# 2. Map the plotting function over the list
# We use 'map' to generate a list of ggplot objects
C_k_plots <- bias_ICC_C_K %>%
  group_by(icc, k_category, e_category) %>%
  group_split() %>%
  # 2. Map over the split data
  map(~ {
    # Extract values for a dynamic and descriptive title
    icc_val <- unique(.x$icc)
    k_val   <- unique(.x$k_category)
    e_val   <- unique(.x$e_category)
    
    # Generate the plot using your generalized function
    plot_simulation_bias(
      data = .x,
      x_var = n_raters,
      y_var = bias,
      col_var = method,
      facet_row = target_icc,
      facet_col = n_objects,
      title = paste0("ICC Type: ", icc_val, " | K = ", k_val, ", Equal = ", e_val)
    )
  })
```


```


## 2.2 Precision: Standard Error  (SD of the simulation distribution) 
We can assess the precision of the generalized ICC compared to the traditional ICC and AOV based ICC by comparing the standard deviation of the simulation distribution

```{r}

```



## 2.3 Discussion: Root Mean Square Error (RMSE)
```{r}
# uniform formatting
my_geom <- list(
  geom_point(position = position_dodge(width = 0.3)),
  geom_line(position = position_dodge(width = 0.3)), 
  geom_errorbar( aes( ymin = rmse - 1.96*rmse_mcse,
                      ymax = rmse + 1.96*rmse_mcse ),
                 position = position_dodge( width=0.3),
                 width = 0 ), 
  geom_hline( yintercept = 0 ),
  theme_minimal() 
)

bias_long_k3_equal <- bias_full |>
  filter(k_category == 3 & e_category == "TRUE") 

p13 <- ggplot(bias_long_k3_equal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K = 3, equal") + 
  my_geom

bias_long_k3_notequal <- bias_full |>
 filter(k_category == 3 & e_category == "FALSE") 

p14 <- ggplot(bias_long_k3_notequal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
 ggtitle("K=3, not equal") + 
 my_geom

bias_long_k5_equal <- bias_full |>
filter(k_category == 5 & e_category == "TRUE") 

p15 <- ggplot(bias_long_k5_equal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
 ggtitle("K = 5, equal") +
 my_geom

bias_long_k5_notequal <- bias_full |>
 filter(k_category == 5 & e_category == "FALSE") 

p16 <- ggplot(bias_long_k5_notequal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K=5, not equal") +
  my_geom

bias_long_k7_equal <- bias_full |>
 filter(k_category == 7 & e_category == "TRUE") 

p17 <- ggplot(bias_long_k7_equal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K=7, equal") +
  my_geom

bias_long_k7_notequal <- bias_full |>
 filter(k_category == 7 & e_category == "FALSE") 

p18 <- ggplot(bias_long_k7_notequal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K=7, not equal") + 
  my_geom

p13
p14
p15
p16
p17
p18
```


# 3. Performance Measures for Standard Errors
- Standard Errors (i.e., standard deviation of replication distribution) 

"A first order question is whether our estimated standard errors are calibrated, meaning whether they are the right size, on average. For this question we will examine the bias of squared standard error estimators. A second order question is whether our estimated standard errors are stable, meaning that they do not vary too much from one dataset to another. For this question we evaluate estimated standard errors in terms of effective degrees of freedom or the standard error (or RMSE) of the standard error."

## 3.1 Calibration

relative bias (calibration) of v  = V_bar (the average variance of across replications, of the variance/SE for each replication utilizing a variance estimator of ICC ) / standard devaition( of T across all replications!)

