---
title: "Ordinal Analysis"
author: Simmons & Girard
date: March 18, 2026
theme: lumen
format:
  html:
    embed-resources: true
---


# 1. Prep Data

## 1.1. Libraries
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(simhelpers)
library(knitr)
library(kableExtra)
```

THIS ANALYZES DATA PRIOR TO SIMULATION RE-RUN (Simulation Universe #2)

## 1.2. Directory & Loading Data

```{r}
DIR <- "~/Documents/Github/MastersThesis/iccs/data"
dat_nest <- readRDS(paste0(DIR,'/Ordinal_run1_full.rds'))
dat_aov_res <- readRDS(paste0(DIR,'/Oridinal_AOV_full.rds'))

#head(dat_nest) 

#Kable (doesn't render?)
# k1 <- kable(dat_nest, format = "html") %>%
#   kable_styling(bootstrap_options = c("striped","hover")) %>%
#   scroll_box(width = "100%", height = "200%")
# k1 
```

## 1.3. Data Wrangling 
```{r}
## unnest the results' list-column  (by hand; example with "icc" column which is numeric) 

# dat <- dat_nest |> 
#   mutate(result = map(result, as_tibble)) |>
#   unnest(result) |>
#   mutate(
#     icc = map_dbl(icc, \(x) pluck(x, .default= NA_real_))) #example

##  unnest efficiently, let R decide the best atomic data type
dat_all <- dat_nest |> 
  mutate(result = map(result, as_tibble)) |>
  unnest(result) |>
  mutate(across(where(is.list), \(col) {
    map_vec(col, \(x) if(is.null(x)) NA else x)
  }))

dat_aov <- dat_aov_res |> mutate(result = map(result, as_tibble)) |>
  unnest(result) |>
  mutate(across(where(is.list), \(col) {
    map_vec(col, \(x) if(is.null(x)) NA else x)
  }))

#set conditions as factors
conds <- c("n_raters", "n_objects", "k_category", "e_category")
dat_all <- dat_all |> 
  mutate(across(.cols = all_of(conds), .fns = as.factor))

dat_aov <- dat_aov |> 
  mutate(across(.cols = all_of(conds), .fns = as.factor))

# Delete nested object
rm(dat_nest)
rm(dat_aov_res)

# Print
head(dat_all)
head(dat_aov)

```

## 1.4 Wide to Long 

Formatting data from wide to long allows easier use of {simhelpers} tidy function for simulations' performance measures. 

```{r}
# add prefixes 
colnames(dat_all)[c(8,12)] <- paste0("pre_", colnames(dat_all)[c(8,12)])
colnames(dat_aov)[8] <- paste0("pre_", colnames(dat_aov)[8]) 

dat_all <- dat_all |>
  pivot_longer(
    cols = starts_with("pre_"),
    names_to = "method",
    names_prefix = ("pre_"),
    values_to =  "value",
  )

dat_aov <- dat_aov |>
  pivot_longer(
    cols = starts_with("pre_"),
    names_to = "method",
    names_prefix = ("pre_"),
    values_to =  "value",
  )
```


<b> In the ordinal case, the parameter grid of conditions is as follows: </b>

```{r}
design_factors <- list(
n_raters = c(3,12,24),
n_objects = c(10,50,100),
target_icc = c(0.40,0.60,0.80),
k_category = c(3,5,7),
e_category = c(TRUE,FALSE) # 1 = equal category prevalence, #0 = unequal linear decay prevalence 
)
#expand grid
params <- expand_grid( !!!design_factors) 
design_factors
head(params)
```


# 2. Performance Measures for Point Estimators
Here, we consider performance measures for point estimators for each of the 162 ordinal conditions. 

We obtain, for each condition and measure (ICC, Generalized ICC, and AOV_ICC):
- Bias (defined as "Mean absolute error" by Oest & Girard, Miratrix & Pustejovsky)
- Bias (defined by Muthen & Muthen)
- Standard Errors (i.e., standard deviation of replication distribution) 
- Root Mean Squared Error (RMSE) 

```{r}
bias_all <- dat_all |>
  #group by each condition (For ordinal, 162 total conditions)
  select(!c(seed, filename)) |>
 # group_by(n_raters, n_objects, target_icc, k_category, e_category) |> 
  summarize(
    simhelpers::calc_absolute( 
      estimates = value, true_param = target_icc, 
      criteria = c("bias","stddev","rmse")
    ), 
    .by = c(n_raters, n_objects, target_icc, k_category, e_category, method)
  )
    
head(bias_all)
```


For AOV measures

Here's an alternate using {simhelpers} for bias, SEs, and RMSE 
- adds MCSE (Monte Carlo Standard Errors)
```{r}
bias_aov <- dat_aov |>
  #group by each condition (For ordinal, 162 total conditions)
  select(!c(seed, filename)) |>
 # group_by(n_raters, n_objects, target_icc, k_category, e_category) |> 
  summarize(
    simhelpers::calc_absolute( 
      estimates = value, true_param = target_icc, 
      criteria = c("bias","stddev","rmse"), 
    ), 
    .by = c(n_raters, n_objects, target_icc, k_category, e_category, method)
  )
    
head(bias_aov)
```

Join together
```{r}
bias_full <-  bias_all |> full_join(bias_aov) 
```

## 2.1 Discussion: Point Estimates Bias
Note: all graphs have corresponding Monte Carlo Standard Error confidence intervals!

### 2.1.1 Bias (Mean Absolute Error)
```{r}
# uniform formatting
my_geom <- list(
  geom_point(position = position_dodge(width = 0.3)),
  geom_line(position = position_dodge(width = 0.3)), 
  geom_errorbar( aes( ymin = bias - 1.96*bias_mcse,
                      ymax = bias + 1.96*bias_mcse ),
                 position = position_dodge( width=0.3),
                 width = 0 ), 
  geom_hline( yintercept = 0 ),
  theme_minimal() 
)

bias_long_k3_equal <- bias_full |>
  filter(k_category == 3 & e_category == "TRUE") 

p1 <- ggplot(bias_long_k3_equal, aes(n_raters, bias, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K = 3, equal") + 
  my_geom

bias_long_k3_notequal <- bias_full |>
 filter(k_category == 3 & e_category == "FALSE") 

p2 <- ggplot(bias_long_k3_notequal, aes(n_raters, bias, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
 ggtitle("K=3, not equal") + 
 my_geom

bias_long_k5_equal <- bias_full |>
filter(k_category == 5 & e_category == "TRUE") 

p3 <- ggplot(bias_long_k5_equal, aes(n_raters, bias, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
 ggtitle("K = 5, equal") +
 my_geom

bias_long_k5_notequal <- bias_full |>
 filter(k_category == 5 & e_category == "FALSE") 

p4 <- ggplot(bias_long_k5_notequal, aes(n_raters, bias, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K=5, not equal") +
  my_geom

bias_long_k7_equal <- bias_full |>
 filter(k_category == 7 & e_category == "TRUE") 

p5 <- ggplot(bias_long_k7_equal, aes(n_raters, bias, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K=7, equal") +
  my_geom

bias_long_k7_notequal <- bias_full |>
 filter(k_category == 7 & e_category == "FALSE") 

p6 <- ggplot(bias_long_k7_notequal, aes(n_raters, bias, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K=7, not equal") + 
  my_geom

p1
p2
p3
p4
p5
p6
```

### 2.1.2 Bias (Muthen & Muthen)
- Here, if bias is less than 10% then it is good
```{r}
my_geom2 <- list(
  geom_point(position = position_dodge(width = 0.3)),
  geom_line(position = position_dodge(width = 0.3)), 
  geom_errorbar( aes( ymin = bias2 - 1.96*bias_mcse,
                      ymax = bias2 + 1.96*bias_mcse ),
                 position = position_dodge( width=0.3),
                 width = 0 ), 
  geom_hline( yintercept = 0 ),
  theme_minimal(), 
  geom_hline(yintercept = -0.10, linetype='dotted', col = 'red'),
  scale_y_continuous(labels = scales::label_percent())
)

bias_long_k3_equal <- bias_full |>
  mutate(bias2 = bias/target_icc) |>
  filter(k_category == 3 & e_category == "TRUE") 

p7 <- ggplot(bias_long_k3_equal, aes(n_raters, bias2, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K = 3, equal") + 
  my_geom2

bias_long_k3_notequal <- bias_full |>
  mutate(bias2 = bias/target_icc) |>
 filter(k_category == 3 & e_category == "FALSE") 

p8 <- ggplot(bias_long_k3_notequal, aes(n_raters, bias2, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
 ggtitle("K=3, not equal") + 
 my_geom2

bias_long_k5_equal <- bias_full |>
  mutate(bias2 = bias/target_icc) |>
filter(k_category == 5 & e_category == "TRUE") 

p9 <- ggplot(bias_long_k5_equal, aes(n_raters, bias2, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
 ggtitle("K = 5, equal") +
 my_geom2

bias_long_k5_notequal <- bias_full |>
  mutate(bias2 = bias/target_icc) |>
 filter(k_category == 5 & e_category == "FALSE") 


p10 <- ggplot(bias_long_k5_notequal, aes(n_raters, bias2, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K=5, not equal") +
  my_geom2

bias_long_k7_equal <- bias_full |>
  mutate(bias2 = bias/target_icc) |>
 filter(k_category == 7 & e_category == "TRUE") 

p11 <- ggplot(bias_long_k7_equal, aes(n_raters, bias2, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K=7, equal") +
  my_geom2

bias_long_k7_notequal <- bias_full |>
  mutate(bias2 = bias/target_icc) |>
 filter(k_category == 7 & e_category == "FALSE") 

p12 <- ggplot(bias_long_k7_notequal, aes(n_raters, bias2, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K=7, not equal") + 
  my_geom2

p7
p8
p9
p10
p11
p12
```


## 2.2 Precision: Standard Error  (SD of the simulation distribution) 
We can assess the precision of the generalized ICC compared to the traditional ICC and AOV based ICC by comparing the standard deviation of the simulation distribution

```{r}

```



## 2.3 Discussion: Root Mean Square Error (RMSE)
```{r}
# uniform formatting
my_geom <- list(
  geom_point(position = position_dodge(width = 0.3)),
  geom_line(position = position_dodge(width = 0.3)), 
  geom_errorbar( aes( ymin = rmse - 1.96*rmse_mcse,
                      ymax = rmse + 1.96*rmse_mcse ),
                 position = position_dodge( width=0.3),
                 width = 0 ), 
  geom_hline( yintercept = 0 ),
  theme_minimal() 
)

bias_long_k3_equal <- bias_full |>
  filter(k_category == 3 & e_category == "TRUE") 

p13 <- ggplot(bias_long_k3_equal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K = 3, equal") + 
  my_geom

bias_long_k3_notequal <- bias_full |>
 filter(k_category == 3 & e_category == "FALSE") 

p14 <- ggplot(bias_long_k3_notequal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
 ggtitle("K=3, not equal") + 
 my_geom

bias_long_k5_equal <- bias_full |>
filter(k_category == 5 & e_category == "TRUE") 

p15 <- ggplot(bias_long_k5_equal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
 ggtitle("K = 5, equal") +
 my_geom

bias_long_k5_notequal <- bias_full |>
 filter(k_category == 5 & e_category == "FALSE") 

p16 <- ggplot(bias_long_k5_notequal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K=5, not equal") +
  my_geom

bias_long_k7_equal <- bias_full |>
 filter(k_category == 7 & e_category == "TRUE") 

p17 <- ggplot(bias_long_k7_equal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K=7, equal") +
  my_geom

bias_long_k7_notequal <- bias_full |>
 filter(k_category == 7 & e_category == "FALSE") 

p18 <- ggplot(bias_long_k7_notequal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K=7, not equal") + 
  my_geom

p13
p14
p15
p16
p17
p18
```


# 3. Performance Measures for Standard Errors
- Standard Errors (i.e., standard deviation of replication distribution) 

"A first order question is whether our estimated standard errors are calibrated, meaning whether they are the right size, on average. For this question we will examine the bias of squared standard error estimators. A second order question is whether our estimated standard errors are stable, meaning that they do not vary too much from one dataset to another. For this question we evaluate estimated standard errors in terms of effective degrees of freedom or the standard error (or RMSE) of the standard error."

## 3.1 Calibration

relative bias (calibration) of v  = V_bar (the average variance of across replications, of the variance/SE for each replication utilizing a variance estimator of ICC ) / standard devaition( of T across all replications!)

