---
title: "Binary Analysis"
author: Simmons & Girard
date: April 7, 2026
theme: lumen
format:
  html:
    embed-resources: true
---

THIS ANALYZES DATA PRIOR TO SIMULATION RE-RUN (Simulation Universe #2)

# 1. Prep Data

## 1.1. Libraries
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(simhelpers)
library(knitr)
library(kableExtra)
```

## 1.2. Directory & Loading Data

```{r}
DIR <- "~/Documents/Github/MastersThesis/iccs/data"
dat_nest <- readRDS(paste0(DIR,'/binary_results_0312_01.rds'))
dat_aov_res <- readRDS(paste0(DIR,'/binary_aOV_full.rds')) #aov results

# dat_all <-  dat_nest |> left_join(dat_aov, by = c("n_raters", "n_objects","target_icc","p","seed"))

#head(dat_nest) 

#Kable (doesn't render?)
# k1 <- kable(dat_nest, format = "html") %>%
#   kable_styling(bootstrap_options = c("striped","hover")) %>%
#   scroll_box(width = "100%", height = "200%")
# k1 
```


## 1.3. Data Wrangling 
```{r}
## unnest the results' list-column  (by hand; example with "icc" column which is numeric) 

# dat <- dat_nest |> 
#   mutate(result = map(result, as_tibble)) |>
#   unnest(result) |>
#   mutate(
#     icc = map_dbl(icc, \(x) pluck(x, .default= NA_real_))) #example

##  unnest efficiently, let R decide the best atomic data type
dat_all <- dat_nest |> 
  mutate(result = map(result, as_tibble)) |>
  unnest(result) |>
  #remove columns for now (unnecessary)
  select(!c(approach, observed, expected, adjusted, boot_results, details, call)) |>
  mutate(across(where(is.list), \(col) {
    map_vec(col, \(x) if(is.null(x)) NA else x)
  }))

dat_aov <- dat_aov_res |>
  mutate(result = map(result, as_tibble)) |>
  unnest(result) |>
  #remove columns for now (unnecessary)
 # select(!c(approach, observed, expected, adjusted, boot_results, details, call)) |>
  mutate(across(where(is.list), \(col) {
    map_vec(col, \(x) if(is.null(x)) NA else x)
  }))

# dat <- dat_all |> 
#   left_join(dat_aov, 
#   by = c("n_raters", "n_objects","target_icc","p","seed","filename"))

#set conditions as factors
conds <- c("n_raters", "n_objects", "p")
dat_all <- dat_all |> 
  mutate(across(.cols = all_of(conds), .fns = as.factor))

dat_aov <- dat_aov |> 
  mutate(across(.cols = all_of(conds), .fns = as.factor))



# Delete nested object
rm(dat_nest)
rm(dat_aov_res)

# Print
head(dat_all)
head(dat_aov)

```

<b> In the binary case, the parameter grid of (54) conditions is as follows: </b>

```{r}
design_factors <- list(
n_raters = c(3,12,24),
n_objects = c(10,50,100),
target_icc = c(0.40,0.60,0.80),
p = c(0.5, 0.8)# for the binary case
)  
#expand grid
params <- expand_grid( !!!design_factors) 
design_factors
head(params)
```

## 1.4 Wide to Long 

Formatting data from wide to long allows easier use of {simhelpers} tidy function for simulations' performance measures. 

```{r}
# add prefixes 
colnames(dat_all)[c(7,11)] <- paste0("pre_", colnames(dat_all)[c(7,11)])
colnames(dat_aov)[7] <- paste0("pre_", colnames(dat_aov)[7]) 

dat_all <- dat_all |>
  pivot_longer(
    cols = starts_with("pre_"),
    names_to = "method",
    names_prefix = ("pre_"),
    values_to =  "value",
  )

dat_aov <- dat_aov |>
  pivot_longer(
    cols = starts_with("pre_"),
    names_to = "method",
    names_prefix = ("pre_"),
    values_to =  "value",
  )
```



# 2. Performance Measures for Point Estimators
Here, we consider measures for point estimators for each of the 54 binary conditions

We obtain, for each condition and measure (ICC, Generalized ICC, and AOV_ICC):
- Bias (defined as "Mean absolute error" by Oest & Girard, Miratrix & Pustejovsky)
- Bias (defined by Muthen & Muthen)
- Standard Errors (i.e., standard deviation of replication distribution) 
- Root Mean Squared Error (RMSE) 
For all measures 
```{r}
bias_all <- dat_all |>
  #group by each condition (For ordinal, 162 total conditions)
  select(!c(seed, filename)) |>
 # group_by(n_raters, n_objects, target_icc, k_category, e_category) |> 
  summarize(
    simhelpers::calc_absolute( 
      estimates = value, true_param = target_icc, 
      criteria = c("bias","stddev","rmse")
    ), 
    .by = c(n_raters, n_objects, target_icc, p, method)
  )
    
head(bias_all)
```

For AOV measures 

```{r}
bias_aov <- dat_aov |>
  #group by each condition (For ordinal, 162 total conditions)
  select(!c(seed, filename)) |>
 # group_by(n_raters, n_objects, target_icc, k_category, e_category) |> 
  summarize(
    simhelpers::calc_absolute( 
      estimates = value, true_param = target_icc, 
      criteria = c("bias","stddev","rmse"), 
    ), 
    .by = c(n_raters, n_objects, target_icc, p, method)
  )
    
head(bias_aov)
```

Join together, sharing CaseNumber
```{r}

bias_full <-  bias_all |> full_join(bias_aov) 
```
## 2.1 Discussion: Point Estimates Bias
Note: all graphs have corresponding Monte Carlo Standard Error confidence intervals!

### 2.1.1 Bias (Mean Absolute Error)
```{r}
# uniform formatting
my_geom <- list(
  geom_point(position = position_dodge(width = 0.3)),
  geom_line(position = position_dodge(width = 0.3)), 
  geom_errorbar( aes( ymin = bias - 1.96*bias_mcse,
                      ymax = bias + 1.96*bias_mcse ),
                 position = position_dodge( width=0.3),
                 width = 0 ), 
  geom_hline( yintercept = 0 ),
  theme_minimal() 
)

bias_long_p0.5 <- bias_full |>
  filter(p ==0.5) 

p1 <- ggplot(bias_long_p0.5, aes(n_raters, bias, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("p=0.5") + 
  my_geom

bias_long_p0.8<- bias_full |>
 filter(p==0.8) 

p2 <- ggplot(bias_long_p0.8, aes(n_raters, bias, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
 ggtitle("p==0.8") + 
 my_geom

p1
p2
```

### 2.1.2 Bias (Muthen & Muthen)
- Here, if bias is less than 10% then it is good
```{r}
my_geom2 <- list(
  geom_point(position = position_dodge(width = 0.3)),
  geom_line(position = position_dodge(width = 0.3)), 
  geom_errorbar( aes( ymin = bias2 - 1.96*bias_mcse,
                      ymax = bias2 + 1.96*bias_mcse ),
                 position = position_dodge( width=0.3),
                 width = 0 ), 
  geom_hline( yintercept = 0 ),
  theme_minimal(), 
  geom_hline(yintercept = -0.10, linetype='dotted', col = 'red'),
  scale_y_continuous(labels = scales::label_percent())
)

bias_long_p0.5 <- bias_full |>
  mutate(bias2 = bias/target_icc) |>
  filter(p==0.5) 

p3 <- ggplot(bias_long_p0.5, aes(n_raters, bias2, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("p=0.5") + 
  my_geom2

bias_long_p0.8 <- bias_full |>
  mutate(bias2 = bias/target_icc) |>
 filter(p==0.8) 

p4 <- ggplot(bias_long_p0.8, aes(n_raters, bias2, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
 ggtitle("p==0.5") + 
 my_geom2


p3
p4
```

## 2.2 Discussion: Root Mean Square Error (RMSE)
```{r}
# uniform formatting
my_geom <- list(
  geom_point(position = position_dodge(width = 0.3)),
  geom_line(position = position_dodge(width = 0.3)), 
  geom_errorbar( aes( ymin = rmse - 1.96*rmse_mcse,
                      ymax = rmse + 1.96*rmse_mcse ),
                 position = position_dodge( width=0.3),
                 width = 0 ), 
  geom_hline( yintercept = 0 ),
  theme_minimal() 
)

bias_long_p0.5 <- bias_full |>
  filter(p==0.5) 

p5 <- ggplot(bias_long_p0.5, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("p==0.5") + 
  my_geom

bias_long_p0.8 <- bias_full |>
 filter(p==0.8) 

p6 <- ggplot(bias_long_p0.8, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
 ggtitle("p==0.8") + 
 my_geom


p5
p6
```







