---
title: "Binary Analysis (Universe #3)"
author: Simmons & Girard
date: May 10, 2026
theme: flatly
format:
  html:
    embed-resources: true
title-block-banner: true
css: custom_theme.css
---


# 1. Setup

Libraries 
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(simhelpers)
library(knitr)
library(kableExtra)
library(gt)
library(ggpubr)
```

{gt} styling 
```{r}
# theme for GT (apa style)

my_gt_theme <- function(gt_object){
  gt_object %>%
  opt_table_font(
    font = list(
      google_font(name = "Arial"), # Preferred
      "serif")) %>%
  tab_options(
    table.border.top.style = "none",
    table.border.bottom.style = "solid",
    table.border.bottom.width = px(2),
    column_labels.border.top.style = "solid",
    column_labels.border.top.width = px(2),
    column_labels.border.bottom.style = "solid",
    column_labels.border.bottom.width = px(1),
    table_body.border.bottom.style = "solid",
    table_body.border.bottom.width = px(2)
  ) 
}

```


# 2. Data Wrangling

## 2.1 Directory & Loading Data
```{r}
DIR <-  "~/Documents/Github/MastersThesis/iccs/data/universe_3_5/"

nest_rds_files <- function(data_df, filename_col, directory_path) {
  # 1. Ensure the directory path ends with a slash for file path construction
  dir <- ifelse(grepl("/$", directory_path), directory_path, paste0(directory_path, "/"))
  
  # 2. Create the new nested column
  # We use sym() to handle the column name if passed as a string
  result_df <- data_df %>%
    mutate(result = map(!!sym(filename_col), function(fname) {
      file_path <- paste0(dir, fname)
      
      # Check if file exists to avoid crashing the whole loop
      if (file.exists(file_path)) {
        return(readRDS(file_path))
      } else {
        warning(paste("File not found:", file_path))
        return(NULL)
      }
    }))
  
  return(result_df)
}
```

## 2.2 Data Wrangling 
<b> In the binary case, the parameter grid of conditions is as follows and add each file to its condition </b>

```{r}
# Join each data file to its condition 
design_factors <- list(
  n_raters = c(3,12,24),
  n_objects = c(10,50,100),
  target_icc = c(0.40,0.60,0.80),
  p = c(0.5, 0.8), # for the binary case
  icc_type = c(1,2,3,4)
) 
#expand grid
params <- expand_grid( !!!design_factors) |>
    mutate(
    seed = 03022026 + 17 * 1:n(), #set seed for each row
    condition = 1:n() * 1,
    filename = paste0("binary_UThree5_",condition,"_",seed,".rds")
  )

dat_nest <- nest_rds_files(params, "filename", DIR)

```

Obtain full dataset
```{r}
dat <- dat_nest |> unnest(result)
#dat_aov <- dat_aov |> unnest(result)

# create full dataset 
#dat <- dplyr::bind_rows(dat_all,dat_aov)

# Delete nested object
rm(dat_nest)
#rm(dat_aov)
#rm(dat_all)

# Print
#head(dat_all)

#set conditions as factors 
conds <- c("n_raters", "n_objects", "p", "icc_type", "condition", "method")
dat <- dat |> 
  mutate(across(.cols = all_of(conds), .fns = as.factor))

```

Subset data so that each g_icc() is matched with its corresponding data generating variance hyper-parameter 

```{r}
#| code-fold: true

# for point-estimate analyses
dat_full <- dat |>
  select(-seed, -filename) |>
  filter(
    (icc_type == 1 & icc == "ICC(A,1)") |
    (icc_type == 2 & icc == "ICC(A,k)") |
    (icc_type == 3 & icc == "ICC(C,1)") |
    (icc_type == 4 & icc == "ICC(C,k)"))

# for CAA analysis 
dat_caa <- dat |> 
  select(-seed, -filename) |> 
  filter(is.na(icc)) |> #select only caa measures 
  bind_rows(dat_full) # match ICCs with correct DGP 

#clean-up 
rm(dat)

```


# 2. Descriptives 
First, let's check how many conditions obtained errors

Errors conditional on ICC type and method
```{r}
errors <- dat_full |>
  mutate(
    across(.cols = c("message","warning","error"), .fns = as.logical)) |> 
  mutate( is_invalid = message | warning | error) |>
 #group_by(c(method)) |>
  summarise(
    total_trials = n(),
    invalid_trials = sum(is_invalid, na.rm = TRUE),
    percent_error = mean(is_invalid, na.rm = TRUE) * 100,
    .by = c(method,icc)
  ) |> 
  mutate(method = fct_recode(method, 
  "LMM {lme4}" = "icc",
  "LMM {glmmtmb}" = "icc_glmmtmb",
  "GLMM {ordinal}" = "g_icc", 
  "ANOVA {lm}" = "aov_icc", 
  ))

```

```{r}
# errors %>%
#   gt() %>%
#   #apply apa style formatting
#   my_gt_theme() %>%
#   tab_header(
#     title = md("*Simulation Failure Rates (Binary Data)*"),
#     subtitle = md("**Comparing Across Method and ICC Type**")
#   ) %>%
#   # Format the percentage column (adds the % sign and rounds to 1 decimal)
#   fmt_percent(
#     columns = percent_error,
#     scale_values = FALSE, # Set to FALSE because you already multiplied by 100!
#     decimals = 2
#   ) %>%
#   # Make the column names look professional
#   cols_label(
#     method = "Method",
#     icc = "ICC Type",
#     total_trials = "Total Trials",
#     invalid_trials = "Invalid Trials",
#     percent_error = "Percent Error"
#   ) %>% 
#   data_color(
#     column = icc,
#     target_columns = everything(),
#     method = "factor",
#     palette = "RColorBrewer::Blues"
#   ) #|>
  #gtsave("failure_rates_binary.png")
```

Negative ICCs conditional on ICC type and method 

```{r}
negative_summary <- dat_full |>
    summarise(
      total_trials = n(),
      negative_count = sum(estimate < 0, na.rm = TRUE),
      .by = c(method,icc), 
      negative_pct = mean(estimate < 0, na.rm = TRUE)
  ) |>
    mutate(method = fct_recode(method, 
    "LMM {lme4}" = "icc",
    "LMM {glmmtmb}" = "icc_glmmtmb",
    "GLMM {ordinal}" = "g_icc", 
    "ANOVA {lm}" = "aov_icc", 
  ))
  

# negative_summary %>%
#   gt() %>%
#   my_gt_theme() %>%
#   tab_header(
#     title = "Negative ICC Rates (Binary Data)",
#     subtitle = "Comparing Across Method and ICC-Type"
#   ) %>%
#   # Format the percentage column (adds the % sign and rounds to 1 decimal)
#   fmt_percent(
#     columns = negative_pct,
#     scale_values = FALSE, # Set to FALSE because you already multiplied by 100!
#     decimals = 2
#   ) %>%
#   # Make the column names look professional
#   cols_label(
#     method = "Method",
#     icc = "ICC Type",
#     total_trials = "Total Trials",
#     negative_count = "Trials Negative ICC",
#     negative_pct = "Percent Negative"
#   ) %>% 
#   data_color(
#     column = icc,
#     target_columns = everything(),
#     method = "factor",
#     palette = "RColorBrewer::Blues"
#   ) #|> 
  #gtsave("negative_rates_binary.png")
```

combine descriptives

```{r}
dplot <- left_join(errors, negative_summary) |>
  gt() |>
  my_gt_theme() |> 
  #apa guidlines say do the titles in word
  # tab_header(
  #   title = "Simulation Failure and Negative ICC Rates (Ordinal Data)",
  #   subtitle = "Comparing Across Method and ICC Type"
  # ) |>
  # Format the percentage column (adds the % sign and rounds to 1 decimal)
  fmt_percent(
    columns = c(percent_error, negative_pct),
    scale_values = FALSE,
    decimals = 2
  ) |>
  # Make the column names look professional
  cols_label(
    method = "Method",
    icc = "ICC Type",
    total_trials = "Total N",
    invalid_trials = "Invalid N",
    percent_error = "Percent N",
    negative_count = "Negative ICC (N)",
    negative_pct = "Percent Negative (N)"
  ) |>
  data_color(
    column = icc,
    target_columns = everything(),
    method = "factor",
    palette = "RColorBrewer::Greys"
  # ) |> 
  #  gtsave("all_descriptives_binary_full.png")
```


# 3. Performance Measures for Point Estimators
Here, we consider performance measures for point estimators for each of the 162 ordinal conditions. 

We obtain, for each condition and measure (ICC, Generalized ICC, and AOV_ICC):
- Bias (defined as "Mean absolute error" by Oest & Girard, Miratrix & Pustejovsky)
- Bias (defined by Muthen & Muthen)
- Standard Errors (i.e., standard deviation of replication distribution) 
- Root Mean Squared Error (RMSE) 
```{r}

bias <- dat_full |>
  #mutate(estimate = na_if(estimate,NaN)) |>
  #some NA (4) in g-ICC(C,k) and negative aov_iccs |>
  filter(estimate >= 0 )|>
  filter(method != "icc") |> # Makes it easier to assess LMM-ICCs since so similar
  summarize(
    simhelpers::calc_absolute( 
      estimates = estimate, true_param = target_icc, 
      criteria = c("bias","stddev","rmse")
    ), 
    .by = c(n_raters, n_objects, target_icc, p, icc_type, condition, method, icc)
  ) |> 
  mutate(bias2 = bias/target_icc) |> #Muthen&Muthen
  mutate(bias2_mcse = bias_mcse/target_icc) |> #Muthen&Muthen
  mutate(Method = fct_recode(method, 
    "LMM {lme4}" = "icc",
    "LMM {glmmtmb}" = "icc_glmmtmb",
    "GLMM {ordinal}" = "g_icc", 
    "ANOVA {lm}" = "aov_icc", )
  ) |>
  mutate(Objects = n_objects) |> 
  mutate(`Target ICC` = target_icc) |>
  mutate(`Category Equality` = p)

bias_no_p <- dat_full |>
  #mutate(estimate = na_if(estimate,NaN)) |>
  #some NA (4) in g-ICC(C,k) and negative aov_iccs |>
  filter(estimate >= 0 )|>
  filter(method != "icc") |> # Makes it easier to assess LMM-ICCs since so similar
  summarize(
    simhelpers::calc_absolute( 
      estimates = estimate, true_param = target_icc, 
      criteria = c("bias","stddev","rmse")
    ), 
    .by = c(n_raters, n_objects, target_icc, icc_type, method, icc)
  ) |> 
  mutate(bias2 = bias/target_icc) |> #Muthen&Muthen
  mutate(bias2_mcse = bias_mcse/target_icc) |> #Muthen&Muthen
  mutate(Method = fct_recode(method, 
    "LMM {lme4}" = "icc",
    "LMM {glmmtmb}" = "icc_glmmtmb",
    "GLMM {ordinal}" = "g_icc", 
    "ANOVA {lm}" = "aov_icc", )
  ) |>
  mutate(Objects = n_objects) |> 
  mutate(`Target ICC` = target_icc)
```


# 3 Discussion: Point Estimates Bias
Note: all graphs have corresponding Monte Carlo Standard Error confidence intervals!

Plot functions 
```{r}

my_geom <- list(
  geom_point(position = position_dodge(width = 0.3)),
  geom_line(position = position_dodge(width = 0.3)),
  geom_errorbar(aes( ymin = bias - 1.96*bias_mcse,
                     ymax = bias + 1.96*bias_mcse ),
                position = position_dodge( width=0.3),
                width = 0 ), 
  geom_hline( yintercept = 0, linetype = "dashed" ),
  theme_minimal(), 
  theme(legend.position = "bottom") 
)

my_geom_mm <- list(
  geom_point(position = position_dodge(width = 0.3)),
  geom_line(position = position_dodge(width = 0.3)), 
  geom_errorbar( aes( ymin = bias2 - 1.96*bias2_mcse,
                      ymax = bias2 + 1.96*bias2_mcse ),
                 position = position_dodge( width=0.3),
                 width = 0 ), 
  geom_hline( yintercept = 0, linetype = "dashed"),
  geom_hline(yintercept = -0.10, color = "red", linetype = "dashed" ),
  geom_hline(yintercept = 0.10, color = "red", linetype = "dashed" ),
  theme_minimal(),
  theme(legend.position = "bottom")
)

my_geom_rmse <- list(
  geom_point(position = position_dodge(width = 0.3)),
  geom_line(position = position_dodge(width = 0.3)), 
  geom_errorbar( aes( ymin = rmse - 1.96*rmse_mcse,
                      ymax = rmse + 1.96*rmse_mcse ),
                 position = position_dodge( width=0.3),
                 width = 0 ), 
  geom_hline( yintercept = 0, linetype = "dashed"),
  theme_minimal() 
)

labs_by_type <- list(
labs(y = "Bias (MAE)"), 
labs(y = "Bias (Muthén & Muthén)"), 
labs(y = "Mean RMSE") 
)

plot_simulation_bias <- function(data, x_var, y_var, col_var, facet_row, facet_col, my_geoms, plot_labs, title = "Simulation Bias") {

 
 
 # pd <- position_dodge(width = 0.5)
  ggplot(data, aes(x = {{x_var}}, 
                   y = {{y_var}}, 
                   col = {{col_var}}, 
                   pch = {{col_var}}, 
                   group = {{col_var}})
                   #fill = {{col_var}})
                    ) +
    # boxplot the probability in Method
  #  geom_boxplot(alpha = 0.7) + 
    #facet_wrap(~{{col_var}}) + 
    # facet_grid handles the two factors dynamically
    facet_grid(rows = vars({{facet_row}}), 
               cols = vars({{facet_col}}), 
               labeller = label_both) +
    # your custom geom layer
    my_geoms + 
    labs(title = title,
         x = plot_labs["n_raters"],
         y = plot_labs["bias"],
         color = substitute(col_var),
         shape = substitute(col_var)) +
    theme_bw() # Or your preferred base theme
}

plot_labels <- c(
  n_raters = "Raters", 
  n_objects = "Objects", 
  target_icc = "Target ICC",
  p = "Probability",
  bias = "MAE Bias (+- MCSE)" 
)

plot_aggregated_facet <- function(data, var_name, bias_type, lab_type) {
  
  # 1. Human-readable labels
  clean_label <- case_when(
    var_name == "n_raters"   ~ "Number of Raters",
    var_name == "n_objects"  ~ "Number of Objects",
    var_name == "target_icc" ~ "Target ICC Value",
    var_name == "p" ~ "Probability (p)",
    TRUE                     ~ var_name
  )
  
  # 2. Aggregate including 'icc' in the grouping
  agg_data <- data |>
    group_by(icc, .data[[var_name]], Method) |>
    summarise(mean_bias = mean({{bias_type}}, na.rm = TRUE),
              sd_bias = sd({{bias_type}}, na.rm = TRUE),
              .groups = "drop")
  
  pd <- position_dodge(width = 0.5)

  # 3. Plot with Faceting by ICC
  p <- ggplot(agg_data, 
    aes(x = factor(.data[[var_name]]), y = mean_bias, 
                       group = Method, color = Method)) +
    geom_line(linewidth = 1, position =pd) +
    geom_point(size = 2, position = pd) + 
    geom_hline(yintercept = 0, 
      color="lemonchiffon4", 
      linetype="dashed", size=0.5) + 
    geom_errorbar(
        aes(ymin = mean_bias - sd_bias, ymax = mean_bias + sd_bias),width = 0.2, alpha = 0.4, position = pd) +
    # Facet by ICC type to see all metrics at once
    facet_wrap(~icc, labeller = label_value) + 
    labs(title = paste("Aggregated Effect of", clean_label),
         x = clean_label, 
         y = "Mean Bias (±1 SD)") +
    theme_bw() +
    theme(legend.position = "bottom") +
    labs_by_type[lab_type]  
  
  if (lab_type == 2){ # add threshold line if Muthen & Muthen Bias
  p <- p + geom_hline(yintercept = c(-0.10,0.10), 
             linetype = "dashed", color = "red")
  
  }
  return(p)
}

```


# 3.1 Bias (Mean Absolute Error)
- MAE individual plots not going to be in paper at all
### MAE Individual plots (collapsed over probability )
probability is shown separately in boxplots 
```{r}


mae_plots <- bias_no_p |>
  #group_by(icc, p) |>
  group_by(icc) |>
  group_split() |>
  # 2. Map over the split data
  map(~ {
    # Extract values for a dynamic and descriptive title
    icc_val <- unique(.x$icc)
   # p_val   <- unique(.x$p)
    
    # Generate the plot using your generalized function
    plot_simulation_bias(
      data = .x,
      x_var = n_raters,
      y_var = bias,
      col_var = Method,
      facet_row = `Target ICC`,
      facet_col = Objects,
      my_geoms = my_geom,
      plot_labs = plot_labels,
      title = paste0("Bias Analysis (Averaged Over Probability): ", icc_val)
    )
  })

mae_plots

```

Save MAE plots (all)

```{r}
#Walk through the plots and save them in organized subdirectories
# walk(mae_plots, ~ {
#   # Pull the metadata from the individual plot's data
#   this_data <- .x$data
#   icc_dir <- as.character(unique(this_data$icc))
#   prob_dir <- as.character(unique(this_data$p))
  
#   # Create directory if it doesn't exist
#   if (!dir.exists(icc_dir)) dir.create(icc_dir)
  
#   # file_name <- paste0(icc_dir, "/K", unique(this_data$k_category), 
#   #                     "_E", unique(this_data$e_category), ".png")
#     file_name <- paste0(icc_dir, prob_dir, "/Binary_MAE_", unique(this_data$icc),
#                        ".png")
  
#   ggsave(file_name, plot = .x, width = 12, height = 9)
# })

```

Save MAE plots (supplementals only)

```{r}
#mae_plots <- mae_plots[-1]
#Walk through the plots and save them in organized subdirectories
# output_dir <- "~/Documents/GitHub/MastersThesis/iccs/all_plots/supplementals/"

# # Create the directory if it doesn't exist
# if (!dir.exists(output_dir)) dir.create(output_dir)

# # Walk through the list by its names to save each plot
# walk(mae_plots, ~ {
#   this_data <- .x$data

#   icc_dir <- as.character(unique(this_data$icc))
#   #icc_p <- as.character(unique(this_data$p))
#   file_path <- file.path(output_dir, paste0(icc_dir, "_mae_overP.png"))
  
#   ggsave(
#     filename = file_path, 
#     plot = .x, 
#     width = 12, 
#     height = 8, 
#     dpi = 300
#   )
# })
```

#### Boxplots 
Boxplots

```{r}
plot_icc_boxplots <- function(data, icc_label) {

  # make clean labels 
  clean_label <- case_when(
    icc_label == "n_raters"   ~ "Number of Raters",
    icc_label == "n_objects"  ~ "Number of Objects",
    icc_label == "target_icc" ~ "Target ICC Value",
    icc_label == "p" ~ "Probability (p)",
    TRUE                     ~ icc_label
  )
  ggplot(data, aes(x = factor(p), y = bias2, fill = Method)) +
    # Draw boxplots for each method side-by-side
    # geom_boxplot(outliers = FALSE, outlier.size = 0.5, alpha = 0.7, outlier.shape = NA) 
    geom_boxplot(alpha = 0.7) +
    # Grid by Equality and Number of Objects
    #facet_grid(p ~ n_objects, labeller = label_both) +
    labs(
      title = paste("Bias Analysis (Binary Data): ", icc_label),
      x = "Probability",
      y = "Percent Bias",
      fill = "Method"
    ) +
    theme_bw() +
    theme(legend.position = "bottom") +
    geom_hline(yintercept = 0, 
      color="lemonchiffon4", 
      linetype="dashed", size=0.5) +
    geom_hline(yintercept = -0.10, color = "red", linetype = "dashed" ) +
    geom_hline(yintercept = 0.10, color = "red", linetype = "dashed" )
}

mae_boxplots <- bias %>%
  group_by(icc) %>%
  group_split() %>%
  map(~ {
    icc_val <- unique(.x$icc)
    
    plot_icc_boxplots(
      data = .x, 
      icc_label = icc_val
    )
  })
```

```{r}
#Walk through the plots and save them in organized subdirectories
# walk(mae_boxplots, ~ {
#   # Pull the metadata from the individual plot's data
#   this_data <- .x$data
#   icc_dir <- as.character(unique(this_data$icc))
  
#   # Create directory if it doesn't exist
#   if (!dir.exists(icc_dir)) dir.create(icc_dir)
  
#   # file_name <- paste0(icc_dir, "/K", unique(this_data$k_category), 
#   #                     "_E", unique(this_data$e_category), ".png")
#     file_name <- paste0(icc_dir, "/Boxplot_", unique(this_data$icc), 
#                        ".png")
  
#   ggsave(file_name, plot = .x, width = 12, height = 9)
#  })

```

save into supplmentals

```{r}
# output_dir <- "~/Documents/GitHub/MastersThesis/iccs/all_plots/supplementals/"

# # Create the directory if it doesn't exist
# if (!dir.exists(output_dir)) dir.create(output_dir)

# # Walk through the list by its names to save each plot
# walk(mae_boxplots, ~ {
#   this_data <- .x$data

#   icc_dir <- as.character(unique(this_data$icc))
#   #icc_p <- as.character(unique(this_data$p))
#   file_path <- file.path(output_dir, paste0(icc_dir, "_boxplots_onP.png"))
  
#   ggsave(
#     filename = file_path, 
#     plot = .x, 
#     width = 12, 
#     height = 8, 
#     dpi = 300
#   )
# })
```


#### Aggregated plots 

Average over factors

```{r}

plot_aggregated_facet <- function(data, var_name) {
  
  # 1. Human-readable labels
  clean_label <- case_when(
    var_name == "n_raters"   ~ "Number of Raters",
    var_name == "n_objects"  ~ "Number of Objects",
    var_name == "target_icc" ~ "Target ICC Value",
    var_name == "p" ~ "Probability",
    var_name == "e_category" ~ "Equality of Categories",
    TRUE                     ~ var_name
  )
  
  # 2. Aggregate including 'icc' in the grouping
  agg_data <- data %>%
    group_by(icc, .data[[var_name]], Method) %>%
    summarise(mean_bias = mean(bias, na.rm = TRUE),
              sd_bias = sd(bias, na.rm = TRUE),
              mean_mcse = sqrt(mean(bias_mcse^2, na.rm = TRUE)),
              .groups = "drop")
  
  pd <- position_dodge(width = 0.5)

  # 3. Plot with Faceting by ICC
  ggplot(agg_data, aes(x = factor(.data[[var_name]]), y = mean_bias, 
                       group = Method, color = Method)) +
    geom_line(linewidth = 1, position =pd) +
    geom_point(size = 2, position = pd) + 
    geom_hline(yintercept = 0, color="lemonchiffon4", linetype="dashed", size=0.5) + 
    geom_errorbar(aes(ymin = mean_bias - mean_mcse, ymax = mean_bias + mean_mcse), 
                  width = 0.2, alpha = 0.4, position = pd) +
    # Facet by ICC type to see all metrics at once
    facet_wrap(~icc, labeller = label_value) + 
    labs(title = paste("Aggregated Effect of", clean_label),
         x = clean_label, 
         y = "Mean MAE Bias (± Mean MCSE)") +
    theme_bw() +
    theme(legend.position = "bottom")
}
```

```{r}
factors_to_test <- c("n_raters", "n_objects", "target_icc", "p")

# Generate a list of faceted plots
final_agg_plots <- map(set_names(factors_to_test), ~ {
  plot_aggregated_facet(bias, .x)
})
```

Save aggregated plots 

```{r}

# # Define the output directory
# output_dir <- "~/Documents/Github/MastersThesis/iccs/aggregated_plots/"

# # Create the folder if it doesn't already exist
# if (!dir.exists(output_dir)) {
#   dir.create(output_dir)
# }

# 4. Save each plot using iwalk (iteration with names)
# iwalk(final_agg_plots, function(p, factor_name) {
  
#   # Construct a clean filename
#   file_name <- file.path(output_dir, paste0("summary_binary_", factor_name, ".png"))
  
#   # Save with high resolution (300 DPI) for clarity
#   ggsave(
#     filename = file_name, 
#     plot = p, 
#     width = 12,    # Wider to accommodate multiple ICC facets
#     height = 8, 
#     units = "in",
#     dpi = 300
#   )
  
#   message(paste("Saved:", file_name))
# })
```

### 2.1.2 Bias (Muthen & Muthen)
- Here, if bias is less than 10% then it is good
```{r}

mm_labs <- c(
  n_raters = "Raters", 
  n_objects = "Objects", 
  target_icc = "Target ICC",
  n_category = "Category",
  bias = "Percent Bias (+- MCSE) " 
)

mm_plots <- bias_no_p |>
  group_by(icc) |>
  group_split() |>
  # 2. Map over the split data
  map(~ {
    # Extract values for a dynamic and descriptive title
    icc_val <- unique(.x$icc)
    #k_val   <- unique(.x$k_category)
    #e_val   <- unique(.x$e_category)
    
    # Generate the plot using your generalized function
    plot_simulation_bias(
      data = .x,
      x_var = n_raters,
      y_var = bias2,
      col_var = Method,
      facet_row = `Target ICC`,
      facet_col = Objects,
      my_geoms = my_geom_mm,
      plot_labs = mm_labs, 
      paste0("Bias Analysis (Averaged Over Probability): ", icc_val)
    )
  })


mm_plots


```


```{r}
#Walk through the plots and save them in organized subdirectories
# walk(mm_plots, ~ {
#   # Pull the metadata from the individual plot's data
#   this_data <- .x$data
#   icc_dir <- as.character(unique(this_data$icc))
#   prob_dir <- as.character(unique(this_data$p))
  
#   # Create directory if it doesn't exist
#   if (!dir.exists(icc_dir)) dir.create(icc_dir)
  
#   # file_name <- paste0(icc_dir, "/K", unique(this_data$k_category), 
#   #                     "_E", unique(this_data$e_category), ".png")
#     file_name <- paste0(icc_dir, prob_dir, "/Binary_MM_", unique(this_data$icc),
#                        ".png")
  
#   ggsave(file_name, plot = .x, width = 12, height = 9)
# })
```
supplmentals only
```{r}
# mm_plots <- mm_plots[-1] #only three plots
# output_dir <- "~/Documents/GitHub/MastersThesis/iccs/all_plots/supplementals/"

# # Create the directory if it doesn't exist
# if (!dir.exists(output_dir)) dir.create(output_dir)

# # Walk through the list by its names to save each plot
# walk(mm_plots, ~ {
#   this_data <- .x$data

#   icc_dir <- as.character(unique(this_data$icc))
#   #icc_p <- as.character(unique(this_data$p))
#   file_path <- file.path(output_dir, paste0(icc_dir, "_percentbias_overP.png"))
  
#   ggsave(
#     filename = file_path, 
#     plot = .x, 
#     width = 12, 
#     height = 8, 
#     dpi = 300
#   )
# })
```


## 3.4 Root Mean Square Error (RMSE)

### RMSE Individual Plots 
```{r}
rmse_labs <- c(
  n_raters = "Raters", 
  n_objects = "Objects", 
  target_icc = "Target ICC",
  n_category = "Category",
  bias = "RMSE (+- MCSE)" 
)

rmse_plots <- bias_no_p |>
  group_by(icc) |>
  group_split() |>
  # 2. Map over the split data
  map(~ {
    # Extract values for a dynamic and descriptive title
    icc_val <- unique(.x$icc)
    # k_val   <- unique(.x$k_category)
    # e_val   <- unique(.x$e_category)
    
    # Generate the plot using your generalized function
    plot_simulation_bias(
      data = .x,
      x_var = n_raters,
      y_var = rmse,
      col_var = Method,
      facet_row = `Target ICC`,
      facet_col = Objects,
      my_geoms = my_geom_rmse,
      plot_labs = rmse_labs, 
      title = paste0("RMSE: ", icc_val)
    )
  })


rmse_plots
```

Save RMSE plots 
```{r}
#Walk through the plots and save them in organized subdirectories
# walk(rmse_plots, ~ {
#   # Pull the metadata from the individual plot's data
#   this_data <- .x$data
#   icc_dir <- as.character(unique(this_data$icc))
#   #prob_dir <- as.character(unique(this_data$p))
  
#   # Create directory if it doesn't exist
#   if (!dir.exists(icc_dir)) dir.create(icc_dir)
  
#   # file_name <- paste0(icc_dir, "/K", unique(this_data$k_category), 
#   #                     "_E", unique(this_data$e_category), ".png")
#     file_name <- paste0(icc_dir, "/Binary_RMSE_", unique(this_data$icc),
#                        ".png")
  
#   ggsave(file_name, plot = .x, width = 12, height = 9)
# })

```
Save into supplmentals

```{r}
# output_dir <- "~/Documents/GitHub/MastersThesis/iccs/all_plots/supplementals/"

# # Create the directory if it doesn't exist
# if (!dir.exists(output_dir)) dir.create(output_dir)

# # Walk through the list by its names to save each plot
# walk(rmse_plots, ~ {
#   this_data <- .x$data

#   icc_dir <- as.character(unique(this_data$icc))
#   #icc_p <- as.character(unique(this_data$p))
#   file_path <- file.path(output_dir, paste0(icc_dir, "_rmse_overP.png"))
  
#   ggsave(
#     filename = file_path, 
#     plot = .x, 
#     width = 12, 
#     height = 8, 
#     dpi = 300
#   )
# })
```

## 2.2 Precision: Standard Error  (SD of the simulation distribution) 
We can assess the precision of the generalized ICC compared to the traditional ICC and AOV based ICC by comparing the standard deviation of the simulation distribution

```{r}

```



## 2.3 Discussion: Root Mean Square Error (RMSE)
```{r}
# uniform formatting
my_geom <- list(
  geom_point(position = position_dodge(width = 0.3)),
  geom_line(position = position_dodge(width = 0.3)), 
  geom_errorbar( aes( ymin = rmse - 1.96*rmse_mcse,
                      ymax = rmse + 1.96*rmse_mcse ),
                 position = position_dodge( width=0.3),
                 width = 0 ), 
  geom_hline( yintercept = 0 ),
  theme_minimal() 
)

bias_long_k3_equal <- bias_full |>
  filter(k_category == 3 & e_category == "TRUE") 

p13 <- ggplot(bias_long_k3_equal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K = 3, equal") + 
  my_geom

bias_long_k3_notequal <- bias_full |>
 filter(k_category == 3 & e_category == "FALSE") 

p14 <- ggplot(bias_long_k3_notequal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
 ggtitle("K=3, not equal") + 
 my_geom

bias_long_k5_equal <- bias_full |>
filter(k_category == 5 & e_category == "TRUE") 

p15 <- ggplot(bias_long_k5_equal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
 ggtitle("K = 5, equal") +
 my_geom

bias_long_k5_notequal <- bias_full |>
 filter(k_category == 5 & e_category == "FALSE") 

p16 <- ggplot(bias_long_k5_notequal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K=5, not equal") +
  my_geom

bias_long_k7_equal <- bias_full |>
 filter(k_category == 7 & e_category == "TRUE") 

p17 <- ggplot(bias_long_k7_equal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K=7, equal") +
  my_geom

bias_long_k7_notequal <- bias_full |>
 filter(k_category == 7 & e_category == "FALSE") 

p18 <- ggplot(bias_long_k7_notequal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K=7, not equal") + 
  my_geom

p13
p14
p15
p16
p17
p18
```


# 3. Performance Measures for Standard Errors
- Standard Errors (i.e., standard deviation of replication distribution) 

"A first order question is whether our estimated standard errors are calibrated, meaning whether they are the right size, on average. For this question we will examine the bias of squared standard error estimators. A second order question is whether our estimated standard errors are stable, meaning that they do not vary too much from one dataset to another. For this question we evaluate estimated standard errors in terms of effective degrees of freedom or the standard error (or RMSE) of the standard error."

## 3.1 Calibration

relative bias (calibration) of v  = V_bar (the average variance of across replications, of the variance/SE for each replication utilizing a variance estimator of ICC ) / standard devaition( of T across all replications!)

## Comparison to CAA

```{r}
dat_avg <- dat_caa |>
  #some NA (4) in g-ICC(C,k) and negative aov_iccs |>
  filter(estimate >= 0 )|>
  filter(method != "icc") |> # Makes it easier to assess LMM since so similar 
  summarize(
    mean_estimate = mean(estimate), 
    .by = c(n_raters, n_objects, target_icc, icc, p, icc_type, condition, method)
  ) |> 
  mutate(
    # 1. Convert double to factor, 2. Rename the levels
    icc_type = as.factor(icc_type),
    icc_type = fct_recode(icc_type,
      "ICC(A,1)" = "1",
      "ICC(A,k)"  = "2",
      "ICC(C,1)" = "3",
      "ICC(C,k)" ="4"
    )
  )
rm(dat_caa) # save RAM space
```

Plot functions

```{r}

plot_multi_facet_correlation <- function(data, 
                                        # col_factor,
                                        caa_metric = "kappa_adj", 
                                        icc_metric = "g_icc"
                                         ) {
  
  # 1. Clean labels for the columns and rows
  get_label <- function(x) {
    case_when(
      x == "n_raters"   ~ "Number of Raters",
      x == "n_objects"  ~ "Number of Objects",
      x == "target_icc" ~ "Target ICC Value",
      x == "k_category" ~ "Number of Categories (K)",
      x == "icc_type"   ~ "ICC Metric Type",
      x == "kappa_adj" ~ "Kappa Adjusted",
      x == "s_bp_adj"  ~ "Brennen-Prediger Adjusted",
      x == "g_icc" ~ "G-ICC",   
      TRUE              ~ x
    )
  }

  # 2. The Join Algorithm (Ensures Kappa and ICC are paired correctly)
  caa_df <- data %>%
    filter(method == caa_metric) %>%
    rename(baseline_val = mean_estimate) %>%
    select(-method)

  icc_df <- data %>%
    filter(method != caa_metric) %>% # Assumes other metrics are your ICCs
    filter(method == icc_metric) %>%
    rename(icc_val = mean_estimate) %>%
    select(-method,icc)

  cor_data <- inner_join(caa_df, icc_df, 
                         by = c("n_raters", "n_objects", "target_icc", "p", "icc_type","condition")) 
                         #%>%
    # Optional: Integrated Outlier Filter (adjust threshold as needed)
    #filter(abs(icc_val) <= 2, abs(baseline_val) <= 2) #do this for now

  # 3. Plotting with Dynamic Faceting
  ggplot(cor_data, aes(x = baseline_val, y = icc_val, color = icc_type)) +
    geom_point(alpha = 0.3, size = 1) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 0.8) +
    stat_cor(aes(label = ..r.label..), method = "pearson", size = 3) +
    # Dynamic Grid: Rows are typically ICC types, Columns are your target factor
    # facet_grid(rows = vars(.data[[row_factor]]), 
    #            cols = vars(.data[[col_factor]]), 
    #            labeller = label_both) +
    labs(
      title = paste("Correlation Analysis:", get_label(caa_metric), "vs", get_label(icc_metric)),
      # subtitle = paste("Faceted by", get_label(row_factor), "and", get_label(col_factor)),
      subtitle = paste("Faceted by ICC Data Generating Process"),
      x = paste("Average", get_label(caa_metric)),
      y = "Average G-ICC Estimate",
      color = "ICC-DGP Type"
    ) +
    theme_bw() +
    theme(legend.position = "bottom") + 
    geom_abline(intercept = 0, slope = 1, color = "orange", linetype = "dashed")
}
```


Create plot


```{r}
# Define the factors you want to loop through
factors_to_analyze <- c("kappa_adj", "s_bp_adj")

# Generate the plots
hyper_cor_plots <- map(set_names(factors_to_analyze), ~ {
  plot_multi_facet_correlation(
    data = dat_avg, 
    caa_metric = .x,
  )
})



# Save the architecture
# output_dir <- "icc_kappa_correlations"
# if (!dir.exists(output_dir)) dir.create(output_dir)

# iwalk(hyper_cor_plots, function(p, factor_name) {
#   ggsave(
#     file.path(output_dir, paste0("hyper_corr_", factor_name, ".png")), 
#     plot = p, width = 16, height = 12, dpi = 300
#   )
# })
```