---
title: "Binary Analysis (Universe #3)"
author: Simmons & Girard
date: May 10, 2026
theme: flatly
format:
  html:
    embed-resources: true
title-block-banner: true
css: custom_theme.css
---


# 1. Setup

Libraries 
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(simhelpers)
library(knitr)
library(kableExtra)
library(gt)
```

{gt} styling 
```{r}
# theme for GT (apa style)

my_gt_theme <- function(gt_object){
  gt_object %>%
  opt_table_font(
    font = list(
      google_font(name = "Arial"), # Preferred
      "serif")) %>%
  tab_options(
    table.border.top.style = "none",
    table.border.bottom.style = "solid",
    table.border.bottom.width = px(2),
    column_labels.border.top.style = "solid",
    column_labels.border.top.width = px(2),
    column_labels.border.bottom.style = "solid",
    column_labels.border.bottom.width = px(1),
    table_body.border.bottom.style = "solid",
    table_body.border.bottom.width = px(2)
  ) 
}

```


# 2. Data Wrangling

## 2.1 Directory & Loading Data
```{r}
DIR <-  "~/Documents/Github/MastersThesis/iccs/data/universe_3_5/"

nest_rds_files <- function(data_df, filename_col, directory_path) {
  # 1. Ensure the directory path ends with a slash for file path construction
  dir <- ifelse(grepl("/$", directory_path), directory_path, paste0(directory_path, "/"))
  
  # 2. Create the new nested column
  # We use sym() to handle the column name if passed as a string
  result_df <- data_df %>%
    mutate(result = map(!!sym(filename_col), function(fname) {
      file_path <- paste0(dir, fname)
      
      # Check if file exists to avoid crashing the whole loop
      if (file.exists(file_path)) {
        return(readRDS(file_path))
      } else {
        warning(paste("File not found:", file_path))
        return(NULL)
      }
    }))
  
  return(result_df)
}
```

## 2.2 Data Wrangling 
<b> In the binary case, the parameter grid of conditions is as follows and add each file to its condition </b>

```{r}
# Join each data file to its condition 
design_factors <- list(
  n_raters = c(3,12,24),
  n_objects = c(10,50,100),
  target_icc = c(0.40,0.60,0.80),
  p = c(0.5, 0.8), # for the binary case
  icc_type = c(1,2,3,4)
) 
#expand grid
params <- expand_grid( !!!design_factors) |>
    mutate(
    seed = 03022026 + 17 * 1:n(), #set seed for each row
    condition = 1:n() * 1,
    filename = paste0("binary_UThree5_",condition,"_",seed,".rds")
  )

dat_nest <- nest_rds_files(params, "filename", DIR)

```

Obtain full dataset
```{r}
dat <- dat_nest |> unnest(result)
#dat_aov <- dat_aov |> unnest(result)

# create full dataset 
#dat <- dplyr::bind_rows(dat_all,dat_aov)

# Delete nested object
rm(dat_nest)
#rm(dat_aov)
#rm(dat_all)

# Print
#head(dat_all)

#set conditions as factors 
conds <- c("n_raters", "n_objects", "p", "icc_type", "condition", "method")
dat <- dat |> 
  mutate(across(.cols = all_of(conds), .fns = as.factor))

```

Subset data so that each g_icc() is matched with its corresponding data generating variance hyper-parameter 

```{r}
#| code-fold: true

# for point-estimate analyses
dat_full <- dat |>
  select(-seed, -filename) |>
  filter(
    (icc_type == 1 & icc == "ICC(A,1)") |
    (icc_type == 2 & icc == "ICC(A,k)") |
    (icc_type == 3 & icc == "ICC(C,1)") |
    (icc_type == 4 & icc == "ICC(C,k)"))

# for CAA analysis 
dat_caa <- dat |> 
  select(-seed, -filename) |> 
  filter(is.na(icc)) |> #select only caa measures 
  bind_rows(dat_full) # match ICCs with correct DGP 

#clean-up 
rm(dat)

```


# 2. Descriptives 
First, let's check how many conditions obtained errors

Errors conditional on ICC type and method
```{r}
errors <- dat_full |>
  mutate(
    across(.cols = c("message","warning","error"), .fns = as.logical)) |> 
  mutate( is_invalid = message | warning | error) |>
 #group_by(c(method)) |>
  summarise(
    total_trials = n(),
    invalid_trials = sum(is_invalid, na.rm = TRUE),
    percent_error = mean(is_invalid, na.rm = TRUE) * 100,
    .by = c(method,icc)
  ) |> 
  mutate(method = fct_recode(method, 
  "LMM {lme4}" = "icc",
  "LMM {glmmtmb}" = "icc_glmmtmb",
  "GLMM {ordinal}" = "g_icc", 
  "ANOVA {lm}" = "aov_icc", 
  ))

```

```{r}
errors %>%
  gt() %>%
  #apply apa style formatting
  my_gt_theme() %>%
  tab_header(
    title = md("*Simulation Failure Rates (Binary Data)*"),
    subtitle = md("**Comparing Across Method and ICC Type**")
  ) %>%
  # Format the percentage column (adds the % sign and rounds to 1 decimal)
  fmt_percent(
    columns = percent_error,
    scale_values = FALSE, # Set to FALSE because you already multiplied by 100!
    decimals = 2
  ) %>%
  # Make the column names look professional
  cols_label(
    method = "Method",
    icc = "ICC Type",
    total_trials = "Total Trials",
    invalid_trials = "Invalid Trials",
    percent_error = "Percent Error"
  ) %>% 
  data_color(
    column = icc,
    target_columns = everything(),
    method = "factor",
    palette = "RColorBrewer::Blues"
  ) #|>
  #gtsave("failure_rates_binary.png")
```

Negative ICCs conditional on ICC type and method 

```{r}
negative_summary <- dat_full |>
    summarise(
      total_trials = n(),
      negative_count = sum(estimate < 0, na.rm = TRUE),
      .by = c(method,icc), 
      negative_pct = mean(estimate < 0, na.rm = TRUE)
  ) |>
    mutate(method = fct_recode(method, 
    "LMM {lme4}" = "icc",
    "LMM {glmmtmb}" = "icc_glmmtmb",
    "GLMM {ordinal}" = "g_icc", 
    "ANOVA {lm}" = "aov_icc", 
  ))
  

negative_summary %>%
  gt() %>%
  my_gt_theme() %>%
  tab_header(
    title = "Negative ICC Rates (Binary Data)",
    subtitle = "Comparing Across Method and ICC-Type"
  ) %>%
  # Format the percentage column (adds the % sign and rounds to 1 decimal)
  fmt_percent(
    columns = negative_pct,
    scale_values = FALSE, # Set to FALSE because you already multiplied by 100!
    decimals = 2
  ) %>%
  # Make the column names look professional
  cols_label(
    method = "Method",
    icc = "ICC Type",
    total_trials = "Total Trials",
    negative_count = "Trials Negative ICC",
    negative_pct = "Percent Negative"
  ) %>% 
  data_color(
    column = icc,
    target_columns = everything(),
    method = "factor",
    palette = "RColorBrewer::Blues"
  ) #|> 
  #gtsave("negative_rates_binary.png")
```

# 3. Performance Measures for Point Estimators
Here, we consider performance measures for point estimators for each of the 162 ordinal conditions. 

We obtain, for each condition and measure (ICC, Generalized ICC, and AOV_ICC):
- Bias (defined as "Mean absolute error" by Oest & Girard, Miratrix & Pustejovsky)
- Bias (defined by Muthen & Muthen)
- Standard Errors (i.e., standard deviation of replication distribution) 
- Root Mean Squared Error (RMSE) 
```{r}

bias <- dat_full |>
  #mutate(estimate = na_if(estimate,NaN)) |>
  #some NA (4) in g-ICC(C,k) and negative aov_iccs |>
  filter(estimate >= 0 )|>
  filter(method != "icc") |> # Makes it easier to assess LMM-ICCs since so similar
  summarize(
    simhelpers::calc_absolute( 
      estimates = estimate, true_param = target_icc, 
      criteria = c("bias","stddev","rmse")
    ), 
    .by = c(n_raters, n_objects, target_icc, p, icc_type, condition, method, icc)
  ) |> 
  mutate(bias2 = bias/target_icc) |> #Muthen&Muthen
  mutate(bias2_mcse = bias_mcse/target_icc) |> #Muthen&Muthen
  mutate(Method = fct_recode(method, 
    "LMM {lme4}" = "icc",
    "LMM {glmmtmb}" = "icc_glmmtmb",
    "GLMM {ordinal}" = "g_icc", 
    "ANOVA {lm}" = "aov_icc", )
  ) |>
  mutate(Objects = n_objects) |> 
  mutate(`Target ICC` = target_icc) |>
  mutate(`Category Equality` = p)
```


# 3 Discussion: Point Estimates Bias
Note: all graphs have corresponding Monte Carlo Standard Error confidence intervals!

Plot functions 
```{r}

my_geom <- list(
  geom_point(position = position_dodge(width = 0.3)),
  geom_line(position = position_dodge(width = 0.3)), 
  geom_errorbar( aes( ymin = bias - 1.96*bias_mcse,
                      ymax = bias + 1.96*bias_mcse ),
                 position = position_dodge( width=0.3),
                 width = 0 ), 
  geom_hline( yintercept = 0 ),
  theme_minimal() 
)

my_geom_rmse <- list(
  geom_point(position = position_dodge(width = 0.3)),
  geom_line(position = position_dodge(width = 0.3)), 
  geom_errorbar( aes( ymin = rmse - 1.96*rmse_mcse,
                      ymax = rmse + 1.96*rmse_mcse ),
                 position = position_dodge( width=0.3),
                 width = 0 ), 
  geom_hline( yintercept = 0),
  theme_minimal() 
)

labs_by_type <- list(
labs(y = "Bias (MAE)"), 
labs(y = "Bias (Muthén & Muthén)"), 
labs(y = "Mean RMSE") 
)

plot_simulation_bias <- function(data, x_var, y_var, col_var, facet_row, facet_col, my_geoms, plot_labs, title = "Simulation Bias") {
  
  ggplot(data, aes(x = {{x_var}}, 
                   y = {{y_var}}, 
                   col = {{col_var}}, 
                   pch = {{col_var}}, 
                   group = {{col_var}})) +
    # facet_grid handles the two factors dynamically
    facet_grid(rows = vars({{facet_row}}), 
               cols = vars({{facet_col}}), 
               labeller = label_both) +
    # your custom geom layer
    my_geoms + 
    labs(title = title,
         x = plot_labs["n_raters"],
         y = plot_labs["bias"],
         color = substitute(col_var),
         shape = substitute(col_var)) +
    theme_bw() # Or your preferred base theme
}

plot_labels <- c(
  n_raters = "Raters", 
  n_objects = "Objects", 
  target_icc = "Target ICC",
  p = "Probability",
  bias = "Bias (MAE)" 
)

plot_aggregated_facet <- function(data, var_name, bias_type, lab_type) {
  
  # 1. Human-readable labels
  clean_label <- case_when(
    var_name == "n_raters"   ~ "Number of Raters",
    var_name == "n_objects"  ~ "Number of Objects",
    var_name == "target_icc" ~ "Target ICC Value",
    var_name == "p" ~ "Probability (p)",
    TRUE                     ~ var_name
  )
  
  # 2. Aggregate including 'icc' in the grouping
  agg_data <- data |>
    group_by(icc, .data[[var_name]], Method) |>
    summarise(mean_bias = mean({{bias_type}}, na.rm = TRUE),
              sd_bias = sd({{bias_type}}, na.rm = TRUE),
              .groups = "drop")
  
  pd <- position_dodge(width = 0.5)

  # 3. Plot with Faceting by ICC
  p <- ggplot(agg_data, 
    aes(x = factor(.data[[var_name]]), y = mean_bias, 
                       group = Method, color = Method)) +
    geom_line(linewidth = 1, position =pd) +
    geom_point(size = 2, position = pd) + 
    geom_hline(yintercept = 0, 
      color="lemonchiffon4", 
      linetype="dashed", size=0.5) + 
    geom_errorbar(
        aes(ymin = mean_bias - sd_bias, ymax = mean_bias + sd_bias),width = 0.2, alpha = 0.4, position = pd) +
    # Facet by ICC type to see all metrics at once
    facet_wrap(~icc, labeller = label_value) + 
    labs(title = paste("Aggregated Effect of", clean_label),
         x = clean_label, 
         y = "Mean Bias (±1 SD)") +
    theme_bw() +
    theme(legend.position = "bottom") +
    labs_by_type[lab_type]  
  
  if (lab_type == 2){ # add threshold line if Muthen & Muthen Bias
  p <- p + geom_hline(yintercept = c(-0.10,0.10), 
             linetype = "dashed", color = "red")
  
  }
  return(p)
}

```


# 3.1 Bias (Mean Absolute Error)

### MAE Individual plots 
```{r}

mae_plots <- bias |>
  group_by(icc, p) |>
  group_split() |>
  # 2. Map over the split data
  map(~ {
    # Extract values for a dynamic and descriptive title
    icc_val <- unique(.x$icc)
    p_val   <- unique(.x$p)
    
    # Generate the plot using your generalized function
    plot_simulation_bias(
      data = .x,
      x_var = n_raters,
      y_var = bias,
      col_var = Method,
      facet_row = `Target ICC`,
      facet_col = Objects,
      my_geoms = my_geom,
      plot_labs = plot_labels,
      title = paste0("ICC Type: ", icc_val, " | Probability = ", p_val)
    )
  })

mae_plots

```

Save MAE plots 

```{r}
#Walk through the plots and save them in organized subdirectories
walk(mae_plots, ~ {
  # Pull the metadata from the individual plot's data
  this_data <- .x$data
  icc_dir <- as.character(unique(this_data$icc))
  prob_dir <- as.character(unique(this_data$p))
  
  # Create directory if it doesn't exist
  if (!dir.exists(icc_dir)) dir.create(icc_dir)
  
  # file_name <- paste0(icc_dir, "/K", unique(this_data$k_category), 
  #                     "_E", unique(this_data$e_category), ".png")
    file_name <- paste0(icc_dir, prob_dir, "/Binary_MAE_", unique(this_data$icc),unique(this_data$p), 
                       ".png")
  
  ggsave(file_name, plot = .x, width = 12, height = 9)
})

```

#### Boxplots 
Boxplots

```{r}
plot_icc_boxplots <- function(data, icc_label) {
  ggplot(data, aes(x = factor(k_category), y = bias, fill = method)) +
    # Draw boxplots for each method side-by-side
    geom_boxplot(outliers = FALSE, outlier.size = 0.5, alpha = 0.7, outlier.shape = NA) +
    # Grid by Equality and Number of Objects
    facet_grid(e_category ~ n_objects, labeller = label_both) +
    labs(
      title = paste("Bias Analysis for ICC Type:", icc_label),
      x = "Number of Categories (K)",
      y = "Bias",
      fill = "Estimation Method"
    ) +
    theme_bw() +
    theme(legend.position = "bottom")
}

all_boxplots <- bias %>%
  group_by(icc) %>%
  group_split() %>%
  map(~ {
    icc_val <- unique(.x$icc)
    
    plot_icc_boxplots(
      data = .x, 
      icc_label = icc_val
    )
  })
```

```{r}
# Walk through the plots and save them in organized subdirectories
walk(all_boxplots, ~ {
  # Pull the metadata from the individual plot's data
  this_data <- .x$data
  icc_dir <- as.character(unique(this_data$icc))
  
  # Create directory if it doesn't exist
  if (!dir.exists(icc_dir)) dir.create(icc_dir)
  
  # file_name <- paste0(icc_dir, "/K", unique(this_data$k_category), 
  #                     "_E", unique(this_data$e_category), ".png")
    file_name <- paste0(icc_dir, "/Boxplot_", unique(this_data$icc), 
                       ".png")
  
  ggsave(file_name, plot = .x, width = 12, height = 9)
})

```




#### Aggregated plots 

Average over factors

```{r}

plot_aggregated_facet <- function(data, var_name) {
  
  # 1. Human-readable labels
  clean_label <- case_when(
    var_name == "n_raters"   ~ "Number of Raters",
    var_name == "n_objects"  ~ "Number of Objects",
    var_name == "target_icc" ~ "Target ICC Value",
    var_name == "k_category" ~ "Number of Categories (K)",
    var_name == "e_category" ~ "Equality of Categories",
    TRUE                     ~ var_name
  )
  
  # 2. Aggregate including 'icc' in the grouping
  agg_data <- data %>%
    group_by(icc, .data[[var_name]], method) %>%
    summarise(mean_bias = mean(bias, na.rm = TRUE),
              sd_bias = sd(bias, na.rm = TRUE),
              .groups = "drop")
  
  pd <- position_dodge(width = 0.5)

  # 3. Plot with Faceting by ICC
  ggplot(agg_data, aes(x = factor(.data[[var_name]]), y = mean_bias, 
                       group = method, color = method)) +
    geom_line(linewidth = 1, position =pd) +
    geom_point(size = 2, position = pd) + 
    geom_hline(yintercept = 0, color="lemonchiffon4", linetype="dashed", size=0.5) + 
    geom_errorbar(aes(ymin = mean_bias - sd_bias, ymax = mean_bias + sd_bias), 
                  width = 0.2, alpha = 0.4, position = pd) +
    # Facet by ICC type to see all metrics at once
    facet_wrap(~icc, labeller = label_value) + 
    labs(title = paste("Aggregated Effect of", clean_label),
         x = clean_label, 
         y = "Mean Bias (±1 SD)") +
    theme_bw() +
    theme(legend.position = "bottom")
}
```

```{r}
factors_to_test <- c("n_raters", "n_objects", "target_icc", "k_category", "e_category")

# Generate a list of faceted plots
final_agg_plots <- map(set_names(factors_to_test), ~ {
  plot_aggregated_facet(bias, .x)
})
```

Save aggregated plots 

```{r}

# Define the output directory
output_dir <- "~/Documents/Github/MastersThesis/iccs/aggregated_plots"

# Create the folder if it doesn't already exist
if (!dir.exists(output_dir)) {
  dir.create(output_dir)
}

# 4. Save each plot using iwalk (iteration with names)
iwalk(final_agg_plots, function(p, factor_name) {
  
  # Construct a clean filename
  file_name <- file.path(output_dir, paste0("summary_", factor_name, ".png"))
  
  # Save with high resolution (300 DPI) for clarity
  ggsave(
    filename = file_name, 
    plot = p, 
    width = 12,    # Wider to accommodate multiple ICC facets
    height = 8, 
    units = "in",
    dpi = 300
  )
  
  message(paste("Saved:", file_name))
})
```

### 2.1.2 Bias (Muthen & Muthen)
- Here, if bias is less than 10% then it is good
```{r}
# my_geom2 <- list(
#   geom_point(position = position_dodge(width = 0.3)),
#   geom_line(position = position_dodge(width = 0.3)), 
#   geom_errorbar( aes( ymin = bias2 - 1.96*bias_mcse,
#                       ymax = bias2 + 1.96*bias_mcse ),
#                  position = position_dodge( width=0.3),
#                  width = 0 ), 
#   geom_hline( yintercept = 0 ),
#   theme_minimal(), 
#   geom_hline(yintercept = -0.10, linetype='dotted', col = 'red'),
#   scale_y_continuous(labels = scales::label_percent())
# )

# bias_long_k3_equal <- bias_full |>
#   mutate(bias2 = bias/target_icc) |>
#   filter(k_category == 3 & e_category == "TRUE") 

# p7 <- ggplot(bias_long_k3_equal, aes(n_raters, bias2, 
#                        col=method, pch=method, group=method ) ) +
#   facet_grid( target_icc ~ n_objects, labeller = label_both ) +
#   ggtitle("K = 3, equal") + 
#   my_geom2

# bias_long_k3_notequal <- bias_full |>
#   mutate(bias2 = bias/target_icc) |>
#  filter(k_category == 3 & e_category == "FALSE") 

# p8 <- ggplot(bias_long_k3_notequal, aes(n_raters, bias2, 
#                        col=method, pch=method, group=method ) ) +
#   facet_grid( target_icc ~ n_objects, labeller = label_both ) +
#  ggtitle("K=3, not equal") + 
#  my_geom2

# bias_long_k5_equal <- bias_full |>
#   mutate(bias2 = bias/target_icc) |>
# filter(k_category == 5 & e_category == "TRUE") 

# p9 <- ggplot(bias_long_k5_equal, aes(n_raters, bias2, 
#                        col=method, pch=method, group=method ) ) +
#   facet_grid( target_icc ~ n_objects, labeller = label_both ) +
#  ggtitle("K = 5, equal") +
#  my_geom2

# bias_long_k5_notequal <- bias_full |>
#   mutate(bias2 = bias/target_icc) |>
#  filter(k_category == 5 & e_category == "FALSE") 


# p10 <- ggplot(bias_long_k5_notequal, aes(n_raters, bias2, 
#                        col=method, pch=method, group=method ) ) +
#   facet_grid( target_icc ~ n_objects, labeller = label_both ) +
#   ggtitle("K=5, not equal") +
#   my_geom2

# bias_long_k7_equal <- bias_full |>
#   mutate(bias2 = bias/target_icc) |>
#  filter(k_category == 7 & e_category == "TRUE") 

# p11 <- ggplot(bias_long_k7_equal, aes(n_raters, bias2, 
#                        col=method, pch=method, group=method ) ) +
#   facet_grid( target_icc ~ n_objects, labeller = label_both ) +
#   ggtitle("K=7, equal") +
#   my_geom2

# bias_long_k7_notequal <- bias_full |>
#   mutate(bias2 = bias/target_icc) |>
#  filter(k_category == 7 & e_category == "FALSE") 

# p12 <- ggplot(bias_long_k7_notequal, aes(n_raters, bias2, 
#                        col=method, pch=method, group=method ) ) +
#   facet_grid( target_icc ~ n_objects, labeller = label_both ) +
#   ggtitle("K=7, not equal") + 
#   my_geom2

# p7
# p8
# p9
# p10
# p11
# p12
```


## 2.2 Precision: Standard Error  (SD of the simulation distribution) 
We can assess the precision of the generalized ICC compared to the traditional ICC and AOV based ICC by comparing the standard deviation of the simulation distribution

```{r}

```



## 2.3 Discussion: Root Mean Square Error (RMSE)
```{r}
# uniform formatting
my_geom <- list(
  geom_point(position = position_dodge(width = 0.3)),
  geom_line(position = position_dodge(width = 0.3)), 
  geom_errorbar( aes( ymin = rmse - 1.96*rmse_mcse,
                      ymax = rmse + 1.96*rmse_mcse ),
                 position = position_dodge( width=0.3),
                 width = 0 ), 
  geom_hline( yintercept = 0 ),
  theme_minimal() 
)

bias_long_k3_equal <- bias_full |>
  filter(k_category == 3 & e_category == "TRUE") 

p13 <- ggplot(bias_long_k3_equal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K = 3, equal") + 
  my_geom

bias_long_k3_notequal <- bias_full |>
 filter(k_category == 3 & e_category == "FALSE") 

p14 <- ggplot(bias_long_k3_notequal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
 ggtitle("K=3, not equal") + 
 my_geom

bias_long_k5_equal <- bias_full |>
filter(k_category == 5 & e_category == "TRUE") 

p15 <- ggplot(bias_long_k5_equal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
 ggtitle("K = 5, equal") +
 my_geom

bias_long_k5_notequal <- bias_full |>
 filter(k_category == 5 & e_category == "FALSE") 

p16 <- ggplot(bias_long_k5_notequal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K=5, not equal") +
  my_geom

bias_long_k7_equal <- bias_full |>
 filter(k_category == 7 & e_category == "TRUE") 

p17 <- ggplot(bias_long_k7_equal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K=7, equal") +
  my_geom

bias_long_k7_notequal <- bias_full |>
 filter(k_category == 7 & e_category == "FALSE") 

p18 <- ggplot(bias_long_k7_notequal, aes(n_raters, rmse, 
                       col=method, pch=method, group=method ) ) +
  facet_grid( target_icc ~ n_objects, labeller = label_both ) +
  ggtitle("K=7, not equal") + 
  my_geom

p13
p14
p15
p16
p17
p18
```


# 3. Performance Measures for Standard Errors
- Standard Errors (i.e., standard deviation of replication distribution) 

"A first order question is whether our estimated standard errors are calibrated, meaning whether they are the right size, on average. For this question we will examine the bias of squared standard error estimators. A second order question is whether our estimated standard errors are stable, meaning that they do not vary too much from one dataset to another. For this question we evaluate estimated standard errors in terms of effective degrees of freedom or the standard error (or RMSE) of the standard error."

## 3.1 Calibration

relative bias (calibration) of v  = V_bar (the average variance of across replications, of the variance/SE for each replication utilizing a variance estimator of ICC ) / standard devaition( of T across all replications!)

