---
title: "Emotional Granularity IRR"
author: Hoemann, Simmons, & Girard
date: May 29, 2026
theme: flatly
format:
  html:
    embed-resources: true
---

# Setup

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(simhelpers)
library(gt)
library(paletteer)
library(ggpubr)
library(gtExtras)
library(vardel)
library(purrr)
```

# Load data 

```{r}
dat <- read_csv("~/Documents/Github/MastersThesis/iccs/data/fullsets/IRR_Emotional_Granularity.csv")

```

# Data Wrangling

```{r}
dat2 <- dat |>
  select(-1) |> # Drop the first column
  mutate(PPID= 1:n()) |> # Change PPID to numbers
  rename(RaterID = PPID) |> #rename
  pivot_longer(
    cols = 2:273,  #everything except the first column
    names_to = "ObjectID", 
    values_to = "Score"
    ) |> 
    separate_wider_delim(
        cols = "ObjectID",
        delim = "_",
        names = c("EmotionID", "ImageID")
    )

dat2
```

# Analysis 
Need to obtain ICC per rater, such that for each rater
- 'rater' = emotion
- 'object' = image id

## Using the AOV framework
```{r, warning = FALSE, message = FALSE}

emotions_aov <- dat2 %>%
    filter(ImageID != "P") |> # don't use practice items |>
    rename(ObjectID = ImageID) %>% # rename to work with functions in {vardel}
    rename(Participants = RaterID) %>% # rename to work with functions in {vardel}
    rename(RaterID = EmotionID) %>% # rename to work with functions in {vardel}
    group_by(Participants) %>%
    group_modify(~ calc_aov_icc(.x))
emotions_aov
```

## Using the g_ICC framework
```{r, warning = FALSE, message = FALSE}
emotions_gicc <- dat2 %>%
    filter(ImageID != "P") |> # don't use practice items |>
    rename(ObjectID = ImageID) %>% # rename to work with functions in {vardel}
    rename(Participants = RaterID) %>% # rename to work with functions in {vardel}
    rename(RaterID = EmotionID) %>% # rename to work with functions in {vardel}
    group_by(Participants) %>%
    group_modify(~ calc_g_ordinal_icc(.x))

emotions_gicc
```



# Comparing Frameworks

First combine all the results 

```{r}
emotions_full <- bind_rows(emotions_aov, emotions_gicc)
emotions_full$method <- as.factor(emotions_full$method)
```

Compare via boxplots
```{r}
ggplot(emotions_full, aes(x = method, y = estimate, fill = method)) +
  geom_boxplot(outlier.alpha = 0.5) +
  theme_bw() + 
  facet_wrap(~ icc, scales = "free_x") + 
  coord_cartesian(ylim = c(-.20, 1)) + 
  geom_hline(yintercept = 0, color = "lemonchiffon4", linetype = "dashed", linewidth = 0.5) +
  labs(
    title = "Distribution of ICCs Across Participants",
    x = "Rater ID",
    y = "ICCs"
  ) +
  theme(
   # axis.text.x = element_text(angle = 45, hjust = 1, size = 8), 
    legend.position = "none" 
  )
```

# Negative ICCs
We see negative ICCs in ANOVA framework

```{r}
negative_iccs <- emotions_full |> 
    filter(estimate< 0)
filt_participants <- unique(negative_iccs$Participants)
```

These same participants do not have negative ICCs in the g-ICC framework 

```{r}
emotions_full |> 
    filter(Participants %in% filt_participants) |>
    select(Participants,method,icc,estimate) |>
    print(n=56)
```

## Rank Comparisons 

### Rank-Rank Scatter plot
-   Shows us where each participant is ranked by one method to the other method. 
-   The 45 degree line is one-to-one mapping between methods 

```{r}
wide_emotions_full <- emotions_full |> 
    group_by(method, icc) |> 
    mutate(Rank = rank(estimate)) |>
    ungroup() |>
    select(Participants, method, icc, Rank) |> 
    pivot_wider(
        names_from = method, 
        values_from = Rank, 
        names_prefix = "Rank_"
    )
wide_emotions_full 
```
<b> Note that the larger ranks have the highest ICC value within method and icc type </b>

Plotting 
```{r}
method_names <- colnames(wide_emotions_full)[3:4]

ggplot(wide_emotions_full, 
aes(
    x = .data[[method_names[1]]], 
    y = .data[[method_names[2]]], 
    color = icc)) +
  geom_point(size = 2.5, alpha = 0.8) +
  # The 45-degree reference line: perfect rank agreement sits exactly on this line
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "grey50", size = 0.8) +
  facet_wrap(~ icc) +
  #scale_x_reverse() + # Optional: Puts Rank 1 at the top/left instead of the bottom
  #scale_y_reverse() + 
  theme_bw() +
  labs(
    title = "Rank Agreement Between Calculation Methods",
    subtitle = "Points far from the dashed line indicate major shifts in rater ranking",
    x = paste("Rater Rank (", gsub("Rank_", "", method_names[1]), ")", sep=""),
    y = paste("Rater Rank (", gsub("Rank_", "", method_names[2]), ")", sep=""),
    color = "ICC Level"
  ) + 
    theme(
        legend.position = "none" 
    )

```

<b> We can conclude that the two methods greatly changed participant ranks </b>
I think this is to be expected, given that the g_icc is less biased. 
For example, some participants obtained negative ICC values in the ANOVA framework, whereas the no participant obtained a negative ICC value in the g_icc framework. 

### Bland-Altman Plot (Tukey mean-difference plot)
This shows us the differences in rank for each participant

```{r}
rank_cols <- colnames(wide_emotions_full)[3:4]
ba_data <- wide_emotions_full %>%
  mutate(
    Rank_Mean = (.data[[rank_cols[1]]] + .data[[rank_cols[2]]]) / 2,
    Rank_Diff = .data[[rank_cols[1]]] - .data[[rank_cols[2]]]
  )
ba_data
```

Plotting 
```{r}
# Calculate overall bias and Limits of Agreement (LoA)
bias <- mean(ba_data$Rank_Diff, na.rm = TRUE)
sd_diff <- sd(ba_data$Rank_Diff, na.rm = TRUE)
upper_loa <- bias + (1.96 * sd_diff)
lower_loa <- bias - (1.96 * sd_diff)

# Create the Bland-Altman Plot faceted by ICC
ggplot(ba_data, aes(x = Rank_Mean, y = Rank_Diff, color = icc)) +
  geom_point(size = 2.5, alpha = 0.8) +
  facet_wrap(~ icc) + 
  
  # Horizontal line at 0 (Perfect Agreement)
  geom_hline(yintercept = 0, linetype = "solid", color = "grey30") +
  
  # Mean Bias Line
  geom_hline(yintercept = bias, linetype = "dashed", color = "blue", size = 0.8) +
  
  # Upper and Lower Limits of Agreement
  geom_hline(yintercept = upper_loa, linetype = "dotted", color = "red", size = 0.8) +
  geom_hline(yintercept = lower_loa, linetype = "dotted", color = "red", size = 0.8) +
  
  facet_wrap(~ icc) +
  theme_minimal() +
  labs(
    title = "Bland-Altman Plot of Participant Rankings",
    subtitle = "Dotted red lines represent the 95% Limits of Agreement",
    x = "Average of the Two Ranks",
    y = "Difference Between the Two Ranks (Method 1 - Method 2)",
    color = "ICC Level"
  ) +
    theme(
        legend.position = "none" 
    )
```
Due to the diamond-shape of the clouds, we can conclude the following: 

    - The participants who have the highest ICC/rank in one method, also have the highest ICC/rank in other other method. 
    - The participants who have the lowest ICC/rank in one method, also have the lowest ICC/rank in the other method
    - The partcipants who had middling ICCs/rank in one method, often get re-ranked in the other method. 

We can pin-point which participants had the largest rank-differnces relative to their rank-mean across methods

```{r}
ba_data |>
    filter(Rank_Diff > upper_loa | Rank_Diff < lower_loa) 
```


### Counter-factual Bland-Altman plot 
To confirm, let's consider a counter-factual Bland-Altman plot if both `aov_icc` and `g_icc` had the same rankings (using the same plot code). 

```{r}
ba_data_counterfactual <- ba_data |>
    group_by(icc) |> 
    mutate(Rank_aov_icc = 1:n(),
    Rank_g_icc = 1:n()) |> 
  mutate(
    Rank_Mean = (.data[[rank_cols[1]]] + .data[[rank_cols[2]]]) / 2,
    Rank_Diff = .data[[rank_cols[1]]] - .data[[rank_cols[2]]]
  )

  # Calculate overall bias and Limits of Agreement (LoA)
bias <- mean(ba_data_counterfactual$Rank_Diff, na.rm = TRUE)
sd_diff <- sd(ba_data_counterfactual$Rank_Diff, na.rm = TRUE)
upper_loa <- bias + (1.96 * sd_diff)
lower_loa <- bias - (1.96 * sd_diff)

# Create the Bland-Altman Plot faceted by ICC
ggplot(ba_data_counterfactual, aes(x = Rank_Mean, y = Rank_Diff, color = icc)) +
  geom_point(size = 2.5, alpha = 0.8) +
  facet_wrap(~ icc) + 
  
  # Horizontal line at 0 (Perfect Agreement)
  geom_hline(yintercept = 0, linetype = "solid", color = "grey30") +
  
  # Mean Bias Line
  geom_hline(yintercept = bias, linetype = "dashed", color = "blue", size = 0.8) +
  
  # Upper and Lower Limits of Agreement
  geom_hline(yintercept = upper_loa, linetype = "dotted", color = "red", size = 0.8) +
  geom_hline(yintercept = lower_loa, linetype = "dotted", color = "red", size = 0.8) +
  
  facet_wrap(~ icc) +
  theme_minimal() +
  labs(
    title = "Counterfactual Bland-Altman Plot of Participant Rankings",
    subtitle = "Dotted red lines represent the 95% Limits of Agreement",
    x = "Average of the Two Ranks",
    y = "Difference Between the Two Ranks (Method 1 - Method 2)",
    color = "ICC Level"
  ) +
    theme(
        legend.position = "none" 
    )
```