---
title: "Ordinal Analysis (Universe #3)"
author: Simmons & Girard
date: May 10, 2026
theme: flatly
format:
  html:
    embed-resources: true
title-block-banner: true
css: custom_theme.css
---

# 1.  Setup 

Libraries 
```{r, warning=FALSE, message=FALSE}
#| code-fold: true

# Load Packages  
library(tidyverse)
library(simhelpers)
library(knitr)
library(kableExtra)
library(gt)
library(paletteer)
library(ggpubr)
# ---
```

{gt} styling template
```{r}
#| code-fold: true
#{gt} Theme --
my_gt_theme <- function(gt_object){
  gt_object |>
  opt_table_font(
    font = list(
      google_font(name = "Arial"), # Preferred
      "serif")) |>
  tab_options(
    table.border.top.style = "none",
    table.border.bottom.style = "solid",
    table.border.bottom.width = px(2),
    column_labels.border.top.style = "solid",
    column_labels.border.top.width = px(2),
    column_labels.border.bottom.style = "solid",
    column_labels.border.bottom.width = px(1),
    table_body.border.bottom.style = "solid",
    table_body.border.bottom.width = px(2)
  ) 
}
# ----
```

# 2. Data Wrangling 

## 2.1 Directory & Loading Data

```{r}
DIR <-  "~/Documents/Github/MastersThesis/iccs/data/universe_3/"
```

```{r}
#| code-fold: true

nest_rds_files <- function(data_df, filename_col, directory_path) {
  # 1. Ensure the directory path ends with a slash for file path construction
  dir <- ifelse(grepl("/$", directory_path), directory_path, paste0(directory_path, "/"))
  
  # 2. Create the new nested column
  # We use sym() to handle the column name if passed as a string
  result_df <- data_df %>%
    mutate(result = map(!!sym(filename_col), function(fname) {
      file_path <- paste0(dir, fname)
      
      # Check if file exists to avoid crashing the whole loop
      if (file.exists(file_path)) {
        return(readRDS(file_path))
      } else {
        warning(paste("File not found:", file_path))
        return(NULL)
      }
    }))
  
  return(result_df)
}
```

## 2.2 Data Wrangling 
<b> In the ordinal case, the parameter grid of conditions is as follows and add each file to its condition </b>

The complete simulations required a post-hoc re-run including the use of {glmmTMB} alongside of {lme4} for linear mixed model fitting. This was obtained in `dat_nest2`

```{r}
#| code-fold: true 

# Paramter Grid
design_factors <- list(
n_raters = c(3,12,24),
n_objects = c(10,50,100),
target_icc = c(0.40,0.60,0.80),
k_category = c(3,5,7),
e_category = c(TRUE,FALSE), 
# 1 = equal category prevalence, #0 = unequal linear decay prevalence 
icc_type = c(1,2,3,4)
)
# Expand grid and
# Join each data file to its condition 
params <- expand_grid( !!!design_factors) |>
    mutate(
    seed = 03122026 + 17 * 1:n(), #set seed for each row,
    condition = 1:n() * 1,
    filename = paste0("ordinal_UThree_",condition,"_",seed,".rds"))

dat_nest <- nest_rds_files(params, "filename", DIR)

params2 <- params |> 
  select(-filename) |>
  mutate(filename = paste0(
    "ordinal_onlyglmmtmb_",condition,"_",seed,".rds"))
  

ODIR <- "~/Documents/Github/MastersThesis/iccs/data/"
#glmmtmb linear fits only 
dat_nest2 <- nest_rds_files(params2, "filename", ODIR)

#head(params)
#head(params2)
```

Here, I  unnested the entire simulation sets to obtain the full data 
```{r}
#| code-fold: true


dat <- dat_nest |> unnest(result)
dat2 <- dat_nest2 |> unnest(result)

# Delete nested object
rm(dat_nest) # no longer neaded 
rm(dat_nest2 ) #no longer needed

# Set each condition of the simulations as factors 
conds <- c("n_raters", "n_objects", "k_category", "icc", "e_category", "icc_type", "condition", "method")
dat <- dat |> 
  mutate(across(.cols = all_of(conds), .fns = as.factor))
dat2 <- dat2 |> 
  mutate(across(.cols = all_of(conds), .fns = as.factor))

```

Critically, datasets were subsetted so that each g_icc() is matched with its corresponding data generating variance (DGP) hyper-parameters

```{r}
#| code-fold: true

# for point-estimate analyses
dat_full <- bind_rows(dat, dat2) |>
  select(-seed, -filename) |>
  filter(
    (icc_type == 1 & icc == "ICC(A,1)") |
    (icc_type == 2 & icc == "ICC(A,k)") |
    (icc_type == 3 & icc == "ICC(C,1)") |
    (icc_type == 4 & icc == "ICC(C,k)"))

# for CAA analysis 
dat_caa <- bind_rows(dat,dat2) |> 
  select(-seed, -filename) |> 
  filter(is.na(icc)) |> #select only caa measures 
  bind_rows(dat_full) # match ICCs with correct DGP 

# Clean up
rm(dat, dat2)

# REDO AOV ICCs due to issue in ICC(A,k) - Jeff found on May 14th

# aov_fixer <- function(khat, n_objects, MSr, MSo, MSe, type) {

#     case_when( 
#       as.character(type) == "ICC(A,1)" ~ (MSo - MSe) / (MSo + ((khat - 1) * MSe) + ((khat / n_objects) * (MSr - MSe))),
#       as.character(type) == "ICC(A,k)" ~ (MSo - MSe) / (MSo + ((MSr - MSe) / n_objects)),
#       as.character(type) == "ICC(C,1)" ~ (MSo - MSe) / (MSo + ((khat - 1) * MSe)), 
#       as.character(type) == "ICC(C,k)" ~ (MSo - MSe) / MSo)
# }

# dat_full_aov <- dat_full |> filter(method == "aov_icc")
# dat_full_not_aov <- dat_full |> filter(method != "aov_icc")

# rm(dat_full) # save space


# dat_full_aov <- dat_full_aov |> 
#   mutate(estimate = aov_fixer(n_raters, n_objects, vr, vs, vsr, icc))

# dat_full <- bind_rows(dat_full_not_aov, dat_full_aov)


```

# 2. Descriptives 

# 2.1 Simulation Error Rates
First, let's check how many conditions obtained errors. Here, I considered a simulation to <i>fail</i> whenever there was a "message", "warning", or "error" associated with any one model fit in {lm} for ANOVA, and {lme4}, {glmmTMB}, and {ordinal} for LMMs. 

Here: Errors are conditional on ICC type and method
```{r}
errors <- dat_full |>
  mutate(
    across(.cols = c("message","warning","error"), .fns = as.logical)) |> 
  mutate( is_invalid = message | warning | error) |>
  summarise(
    total_trials = n(),
    invalid_trials = sum(is_invalid, na.rm = TRUE),
    percent_error = mean(is_invalid, na.rm = TRUE) * 100,
    .by = c(method,icc)
  ) |>
  arrange(match(method, c("icc","icc_glmmtmb", "g_icc","aov_icc"))
  ) |>
  arrange(icc) |> 
  mutate(method = fct_recode(method, 
  "LMM {lme4}" = "icc",
  "LMM {glmmtmb}" = "icc_glmmtmb",
  "GLMM {ordinal}" = "g_icc", 
  "ANOVA {lm}" = "aov_icc", 
  ))

```

Print {gt} Table 
```{r}
errors |>
  gt() |>
  my_gt_theme() |>
  tab_header(
    title = "Simulation Failure Rates (Ordinal Data)",
    subtitle = "Comparing Across Method and ICC-Type"
  ) |>
  # Format the percentage column (adds the % sign and rounds to 1 decimal)
  fmt_percent(
    columns = percent_error,
    scale_values = FALSE,
    decimals = 2
  ) |>
  # Make the column names look professional
  cols_label(
    method = "Method",
    icc = "ICC Type",
    total_trials = "Total Trials",
    invalid_trials = "Invalid Trials",
    percent_error = "Percent Error"
  ) |>
  data_color(
    column = icc,
    target_columns = everything(),
    method = "factor",
    palette = "RColorBrewer::Blues"
  ) #|>
  #gtsave("failure_rates_ordinal.png")
```

# 2.2 Negative ICC rates 
Negative ICCs conditional on ICC type and method 

```{r}
negative_summary <- dat_full |>
  summarise(
    total_trials = n(),
    negative_count = sum(estimate < 0, na.rm = TRUE),
    .by = c(method,icc), 
    negative_pct = mean(estimate < 0, na.rm = TRUE)
  ) |>
  arrange(match(method, c("icc","icc_glmmtmb", "g_icc","aov_icc"))
  ) |>
  arrange(icc) |> 
  mutate(method = fct_recode(method, 
  "LMM {lme4}" = "icc",
  "LMM {glmmtmb}" = "icc_glmmtmb",
  "GLMM {ordinal}" = "g_icc", 
  "ANOVA {lm}" = "aov_icc", 
  ))

negative_summary |>
  gt() |>
  my_gt_theme() |>
  tab_header(
    title = "Negative ICC Rates (Ordinal Data)",
    subtitle = "Comparing Across Method and ICC-Type"
  ) |>
  # Format the percentage column (adds the % sign and rounds to 1 decimal)
  fmt_percent(
    columns = negative_pct,
    scale_values = FALSE, 
    decimals = 2
  ) |>
  # Make the column names look professional
  cols_label(
    method = "Method",
    icc = "ICC Type",
    total_trials = "Total Trials",
    negative_count = "Trials Negative ICC",
    negative_pct = "Percent Negative"
  ) |>
  data_color(
    column = icc,
    target_columns = everything(),
    method = "factor",
    palette = "RColorBrewer::Blues"
  ) |> 
  gtsave("negative_ICC_rates_ordinal.png")
```

# 3. Performance Measures for Point Estimators
Here, we consider performance measures for point estimators for each of the 162 ordinal conditions. 

We obtain, for each condition and measure (ICC, Generalized ICC, and AOV_ICC):
- Bias (defined as "Mean absolute error" by Oest & Girard, Miratrix & Pustejovsky)
- Bias (defined by Muthen & Muthen)
- Standard Errors (i.e., standard deviation of replication distribution) 
- Root Mean Squared Error (RMSE) 
```{r}

bias <- dat_full |>
  #some NA (4) in g-ICC(C,k) and negative aov_iccs |>
  filter(estimate >= 0 )|>
  filter(method != "icc") |> # Makes it easier to assess LMM since so similar 
  summarize(
    simhelpers::calc_absolute( 
      estimates = estimate, true_param = target_icc, 
      criteria = c("bias","stddev","rmse")
    ), 
    .by = c(n_raters, n_objects, target_icc, icc, k_category, e_category, icc_type, condition, method)
  ) |>
  mutate(bias2 = bias/target_icc) |> #Muthen&Muthen
  mutate(bias2_mcse = bias_mcse/target_icc) |> #Muthen&Muthen
  mutate(Method = fct_recode(method, 
    "LMM {lme4}" = "icc",
    "LMM {glmmtmb}" = "icc_glmmtmb",
    "GLMM {ordinal}" = "g_icc", 
    "ANOVA {lm}" = "aov_icc", )
  ) |> 
  mutate(Objects = n_objects) |> 
  mutate(`Target ICC` = target_icc) |>
  mutate(`Category Equality` = e_category)

  
  
```

Plot functions 
```{r}
#| code-fold: true

my_geom <- list(
  geom_point(position = position_dodge(width = 0.3)),
  geom_line(position = position_dodge(width = 0.3)), 
  geom_errorbar( aes( ymin = bias - 1.96*bias_mcse,
                      ymax = bias + 1.96*bias_mcse ),
                 position = position_dodge( width=0.3),
                 width = 0 ), 
  geom_hline( yintercept = 0 ),
  theme_minimal() 
)

my_geom2 <- list(
  geom_point(position = position_dodge(width = 0.3)),
  geom_line(position = position_dodge(width = 0.3)), 
  geom_errorbar( aes( ymin = bias2 - 1.96*bias2_mcse,
                      ymax = bias2 + 1.96*bias2_mcse ),
                 position = position_dodge( width=0.3),
                 width = 0 ), 
  geom_hline( yintercept = 0),
  geom_hline(yintercept = -0.10, color = "red", linetype = "dashed" ),
  geom_hline(yintercept = 0.10, color = "red", linetype = "dashed" ),
  theme_minimal() 
)

my_geom_rmse <- list(
  geom_point(position = position_dodge(width = 0.3)),
  geom_line(position = position_dodge(width = 0.3)), 
  geom_errorbar( aes( ymin = rmse - 1.96*rmse_mcse,
                      ymax = rmse + 1.96*rmse_mcse ),
                 position = position_dodge( width=0.3),
                 width = 0 ), 
  geom_hline( yintercept = 0),
  theme_minimal() 
)

labs_by_type <- list(
labs(y = "Bias (MAE)"), 
labs(y = "Bias (Muthén & Muthén)"), 
labs(y = "Mean RMSE") 
)



plot_simulation_bias <- function(data, x_var, y_var, col_var, facet_row, facet_col, my_geoms, plot_labs, title = "Simulation Bias") {
  
  ggplot(data, aes(x = {{x_var}}, 
                   y = {{y_var}}, 
                   col = {{col_var}}, 
                   pch = {{col_var}}, 
                   group = {{col_var}})) +
    # facet_grid handles the two factors dynamically
    facet_grid(rows = vars({{facet_row}}), 
               cols = vars({{facet_col}}), 
               labeller = label_both) +
    # your custom geom layer
    my_geoms + 
    # labs(title = title,
    #      x = substitute(x_var),
    #      y = substitute(y_var),
    #      color = substitute(col_var),
    #      shape = substitute(col_var)) +
    labs(title = title,
         x = plot_labs["n_raters"],
         y = plot_labs["bias"],
         color = substitute(col_var),
         shape = substitute(col_var)) +
    theme_bw() # Or your preferred base theme
}

plot_labels <- c(
  n_raters = "Raters", 
  n_objects = "Objects", 
  target_icc = "Target ICC",
  n_category = "Category",
  bias = "Bias (MAE)" 
)

plot_aggregated_facet <- function(data, var_name, bias_type, lab_type) {
  
  # 1. Human-readable labels
  clean_label <- case_when(
    var_name == "n_raters"   ~ "Number of Raters",
    var_name == "n_objects"  ~ "Number of Objects",
    var_name == "target_icc" ~ "Target ICC Value",
    var_name == "k_category" ~ "Number of Categories (K)",
    var_name == "e_category" ~ "Equality of Categories",
    TRUE                     ~ var_name
  )
  
  # 2. Aggregate including 'icc' in the grouping
  agg_data <- data |>
    group_by(icc, .data[[var_name]], Method) |>
    summarise(mean_bias = mean({{bias_type}}, na.rm = TRUE),
              sd_bias = sd({{bias_type}}, na.rm = TRUE),
              .groups = "drop")
  
  pd <- position_dodge(width = 0.5)

  # 3. Plot with Faceting by ICC
  p <- ggplot(agg_data, 
    aes(x = factor(.data[[var_name]]), y = mean_bias, 
                       group = Method, color = Method)) +
    geom_line(linewidth = 1, position =pd) +
    geom_point(size = 2, position = pd) + 
    geom_hline(yintercept = 0, 
      color="lemonchiffon4", 
      linetype="dashed", size=0.5) + 
    geom_errorbar(
        aes(ymin = mean_bias - sd_bias, ymax = mean_bias + sd_bias),width = 0.2, alpha = 0.4, position = pd) +
    # Facet by ICC type to see all metrics at once
    facet_wrap(~icc, labeller = label_value) + 
    labs(title = paste("Aggregated Effect of", clean_label),
         x = clean_label, 
         y = "Mean Bias (±1 SD)") +
    theme_bw() +
    theme(legend.position = "bottom") +
    labs_by_type[lab_type]  
  
  if (lab_type == 2){ # add threshold line if Muthen & Muthen Bias
  p <- p + geom_hline(yintercept = c(-0.10,0.10), 
             linetype = "dashed", color = "red")
  
  }
  return(p)
}


```


## 3.1 Bias (Mean Absolute Error)

### MAE Individual plots 
```{r}
mae_plots <- bias |>
  group_by(icc, k_category, e_category) |>
  group_split() |>
  # 2. Map over the split data
  map(~ {
    # Extract values for a dynamic and descriptive title
    icc_val <- unique(.x$icc)
    k_val   <- unique(.x$k_category)
    e_val   <- unique(.x$e_category)
    
    # Generate the plot using your generalized function
    plot_simulation_bias(
      data = .x,
      x_var = n_raters,
      y_var = bias,
      col_var = Method,
      facet_row = `Target ICC`,
      facet_col = Objects,
      my_geoms = my_geom,
      plot_labs = plot_labels,
      title = paste0("ICC Type: ", icc_val, " | K = ", k_val, ", Equal = ", e_val)
    )
  })

mae_plots
```

### MAE Boxplots 
Boxplots

```{r}
plot_icc_boxplots <- function(data, icc_label) {
  ggplot(data, aes(x = factor(k_category), y = bias, fill = Method)) +
    # Draw boxplots for each method side-by-side
    geom_boxplot(outliers = FALSE, outlier.size = 0.5, alpha = 0.7, outlier.shape = NA) +
    # Grid by Equality and Number of Objects
    facet_grid(`Category Equality` ~ Objects, labeller = label_both) +
    labs(
      title = paste("Bias Analysis for ICC Type:", icc_label),
      x = "Number of Categories (K)",
      y = "Bias (MAE)",
      fill = "Estimation Method"
    ) +
    theme_bw() +
    theme(legend.position = "bottom")
}

mae_boxplots <- bias %>%
  group_by(icc) %>%
  group_split() %>%
  map(~ {
    icc_val <- unique(.x$icc)
    
    plot_icc_boxplots(
      data = .x, 
      icc_label = icc_val
    )
  })
```
Save boxplots
```{r}
# # Walk through the plots and save them in organized subdirectories
# walk(all_boxplots, ~ {
#   # Pull the metadata from the individual plot's data
#   this_data <- .x$data
#   icc_dir <- as.character(unique(this_data$icc))
  
#   # Create directory if it doesn't exist
#   if (!dir.exists(icc_dir)) dir.create(icc_dir)
  
#   # file_name <- paste0(icc_dir, "/K", unique(this_data$k_category), 
#   #                     "_E", unique(this_data$e_category), ".png")
#     file_name <- paste0(icc_dir, "/Boxplot_", unique(this_data$icc), 
#                        ".png")
  
#   ggsave(file_name, plot = .x, width = 12, height = 9)
# })

```

### MAE Aggregated plots 

Average over factors
```{r}
 factors_to_test <- c("n_raters", "n_objects", "target_icc", "k_category", "e_category")

#  plot
 mae_aggregate_plots <- map(set_names(factors_to_test), 
 ~ {plot_aggregated_facet(bias, .x, bias, 1)
 })
 mae_aggregate_plots
```

Save aggregated plots 

```{r}

# # Define the output directory
# output_dir <- "~/Documents/Github/MastersThesis/iccs/aggregated_plots"

# # Create the folder if it doesn't already exist
# if (!dir.exists(output_dir)) {
#   dir.create(output_dir)
# }

# # 4. Save each plot using iwalk (iteration with names)
# iwalk(final_agg_plots, function(p, factor_name) {
  
#   # Construct a clean filename
#   file_name <- file.path(output_dir, paste0("summary_", factor_name, ".png"))
  
#   # Save with high resolution (300 DPI) for clarity
#   ggsave(
#     filename = file_name, 
#     plot = p, 
#     width = 12,    # Wider to accommodate multiple ICC facets
#     height = 8, 
#     units = "in",
#     dpi = 300
#   )
  
#   message(paste("Saved:", file_name))
# })
```

## 3.2 Bias (Muthen & Muthen)
- Here, if bias is less than 10% then it is good

### Muthen & Muthen Individual Plots 
```{r}

mm_labs <- c(
  n_raters = "Raters", 
  n_objects = "Objects", 
  target_icc = "Target ICC",
  n_category = "Category",
  bias = "Bias (Muthén & Muthén)" 
)

mm_plots <- bias |>
  group_by(icc, k_category, e_category) |>
  group_split() |>
  # 2. Map over the split data
  map(~ {
    # Extract values for a dynamic and descriptive title
    icc_val <- unique(.x$icc)
    k_val   <- unique(.x$k_category)
    e_val   <- unique(.x$e_category)
    
    # Generate the plot using your generalized function
    plot_simulation_bias(
      data = .x,
      x_var = n_raters,
      y_var = bias2,
      col_var = Method,
      facet_row = `Target ICC`,
      facet_col = Objects,
      my_geoms = my_geom2,
      plot_labs = mm_labs, 
      title = paste0("ICC Type: ", icc_val, " | K = ", k_val, ", Equal = ", e_val)
    )
  })


mm_plots
```

### Muthen & Muthen Aggregated plots 

```{r}
 factors_to_test <- c("n_raters", "n_objects", "target_icc", "k_category", "e_category")

# # Generate a list of faceted plots
 mm_aggregate_plots <- map(set_names(factors_to_test), 
 ~ {plot_aggregated_facet(bias, .x, bias2, 2)
 })

 mm_aggregate_plots

```

## 3.3 Precision: Standard Error  (SD of the simulation distribution)  (Skip for now)
We can assess the precision of the generalized ICC compared to the traditional ICC and AOV based ICC by comparing the standard deviation of the simulation distribution

## 3.4 Root Mean Square Error (RMSE)

### RMSE Individual Plots 
```{r}
rmse_labs <- c(
  n_raters = "Raters", 
  n_objects = "Objects", 
  target_icc = "Target ICC",
  n_category = "Category",
  bias = "RMSE" 
)

rmse_plots <- bias |>
  group_by(icc, k_category, e_category) |>
  group_split() |>
  # 2. Map over the split data
  map(~ {
    # Extract values for a dynamic and descriptive title
    icc_val <- unique(.x$icc)
    k_val   <- unique(.x$k_category)
    e_val   <- unique(.x$e_category)
    
    # Generate the plot using your generalized function
    plot_simulation_bias(
      data = .x,
      x_var = n_raters,
      y_var = rmse,
      col_var = Method,
      facet_row = `Target ICC`,
      facet_col = Objects,
      my_geoms = my_geom_rmse,
      plot_labs = rmse_labs, 
      title = paste0("ICC Type: ", icc_val, " | K = ", k_val, ", Equal = ", e_val)
    )
  })


rmse_plots
```

### RMSE Aggregated plots 
Aggregated RMSE plots 
```{r}
 factors_to_test <- c("n_raters", "n_objects", "target_icc", "k_category", "e_category")

# # Generate a list of faceted plots
 rmse_aggregate_plots <- map(set_names(factors_to_test), 
 ~ {plot_aggregated_facet(bias, .x, rmse, 3)
 })

 rmse_aggregate_plots
```

# 3.4 Performance Measures for Standard Errors (Skip for now)
- Standard Errors (i.e., standard deviation of replication distribution) 

"A first order question is whether our estimated standard errors are calibrated, meaning whether they are the right size, on average. For this question we will examine the bias of squared standard error estimators. A second order question is whether our estimated standard errors are stable, meaning that they do not vary too much from one dataset to another. For this question we evaluate estimated standard errors in terms of effective degrees of freedom or the standard error (or RMSE) of the standard error."

## 3.5 Calibration (Skip for now)

Relative bias (calibration) of v  = V_bar (the average variance of across replications, of the variance/SE for each replication utilizing a variance estimator of ICC ) / standard devaition( of T across all replications!)

# 4. Comparison to CAA


```{r}
dat_avg <- dat_caa |>
  #some NA (4) in g-ICC(C,k) and negative aov_iccs |>
  filter(estimate >= 0 )|>
  filter(method != "icc") |> # Makes it easier to assess LMM since so similar 
  summarize(
    mean_estimate = mean(estimate), 
    .by = c(n_raters, n_objects, target_icc, icc, k_category, e_category, icc_type, condition, method)
  )
rm(dat_caa) # save RAM space
```

Plot functions

```{r}
# plot_hyper_parameter_correlation<- function(data, target_factor, caa_metric = "kappa_adj") {
  
#   # 1. Clean labeling
#   clean_label <- case_when(
#     target_factor == "n_raters"   ~ "Number of Raters",
#     target_factor == "n_objects"  ~ "Number of Objects",
#     target_factor == "target_icc" ~ "Target ICC Value",
#     target_factor == "k_category" ~ "Number of Categories (K)",
#     TRUE                          ~ target_factor
#   )

#   # 2. SEPARATE and JOIN to ensure correct pairing
#   # This prevents the "horizontal line" issue caused by messy pivots
  
#   caa_df <- data %>%
#     filter(method == caa_metric) %>%
#     rename(baseline_val = mean_estimate) %>%
#     select(-method)

#   icc_df <- data %>%
#     filter(method != caa_metric) %>% # Assumes other metrics are your ICCs
#     rename(icc_val = mean_estimate) %>%
#     select(-method,icc)

#   # Join them on all simulation factors to ensure 1:1 matching
#   cor_data <- inner_join(caa_df, icc_df, 
#                          by = c("n_raters", "n_objects", "target_icc", "k_category", "e_category", "icc_type","condition"))

#   # 3. Dynamic Plotting
#   ggplot(cor_data, aes(x = baseline_val, y = icc_val, color = icc_type)) +
#     geom_point(alpha = 0.4, size = 1.2) +
#     geom_smooth(method = "lm", se = FALSE, linewidth = 0.8) +
#     stat_cor(aes(label = ..r.label..), method = "pearson", label.y.npc = "top") +
#     # Use any_of for the dynamic faceting
#     facet_grid(rows = vars(icc_type), cols = vars(any_of(target_factor)), labeller = label_both) +
#     labs(
#       title = paste("Correlation:", baseline_metric, "vs. ICC Types"),
#       subtitle = paste("Faceted by", clean_label),
#       x = paste("Average", baseline_metric),
#       y = "Average ICC Estimate",
#       color = "Estimation Method"
#     ) +
#     theme_bw()
# }
```

```{r}

plot_multi_facet_correlation <- function(data, 
                                         col_factor, 
                                         row_factor = "icc_type", 
                                         caa_metric = "kappa_adj") {
  
  # 1. Clean labels for the columns and rows
  get_label <- function(x) {
    case_when(
      x == "n_raters"   ~ "Number of Raters",
      x == "n_objects"  ~ "Number of Objects",
      x == "target_icc" ~ "Target ICC Value",
      x == "k_category" ~ "Number of Categories (K)",
      x == "icc_type"   ~ "ICC Metric Type",
      TRUE              ~ x
    )
  }

  # 2. The Join Algorithm (Ensures Kappa and ICC are paired correctly)
  caa_df <- data %>%
    filter(method == caa_metric) %>%
    rename(baseline_val = mean_estimate) %>%
    select(-method)

  icc_df <- data %>%
    filter(method != caa_metric) %>% # Assumes other metrics are your ICCs
    filter(method == "g_icc") %>%
    rename(icc_val = mean_estimate) %>%
    select(-method,icc)

  cor_data <- inner_join(caa_df, icc_df, 
                         by = c("n_raters", "n_objects", "target_icc", "k_category", "e_category", "icc_type","condition")) %>%
    # Optional: Integrated Outlier Filter (adjust threshold as needed)
    filter(abs(icc_val) <= 2, abs(baseline_val) <= 2) #do this for now

  # 3. Plotting with Dynamic Faceting
  ggplot(cor_data, aes(x = baseline_val, y = icc_val)) +
    geom_point(alpha = 0.3, size = 1) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 0.8) +
    stat_cor(aes(label = ..r.label..), method = "pearson", size = 3) +
    # Dynamic Grid: Rows are typically ICC types, Columns are your target factor
    facet_grid(rows = vars(.data[[row_factor]]), 
               cols = vars(.data[[col_factor]]), 
               labeller = label_both) +
    labs(
      title = paste("Correlation Analysis:", caa_metric, "vs. G-ICC"),
      subtitle = paste("Faceted by", get_label(row_factor), "and", get_label(col_factor)),
      x = paste("Average", caa_metric),
      y = "Average ICC Estimate"
    ) +
    theme_bw() +
    theme(legend.position = "bottom") + 
    geom_abline(intercept = 0, slope = 1, color = "orange", linetype = "dashed")
}
```


Create plot


```{r}
# Define the factors you want to loop through
factors_to_analyze <- c("n_raters", "n_objects", "k_category")

# Generate the plots
hyper_cor_plots <- map(set_names(factors_to_analyze), ~ {
  plot_multi_facet_correlation(
    data = dat_avg, 
    col_factor = .x,
  )
})



# Save the architecture
# output_dir <- "icc_kappa_correlations"
# if (!dir.exists(output_dir)) dir.create(output_dir)

# iwalk(hyper_cor_plots, function(p, factor_name) {
#   ggsave(
#     file.path(output_dir, paste0("hyper_corr_", factor_name, ".png")), 
#     plot = p, width = 16, height = 12, dpi = 300
#   )
# })
```